---
title: "Location Exploration"
author: "Maddee Wiggins (FlowWest)"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(leaflet)
library(sf)
library(tidyverse)
```

### **Objectives**

1.  download NHD layers for Klamath Basin
-   NHDplus geodatabase downloaded from: <https://www.sciencebase.gov/catalog/item/5d30c29ae4b01d82ce84aa5e>
-   NHDplus user guide: <https://pubs.usgs.gov/of/2019/1096/ofr20191096.pdf>
2.  explore NHD layer and HUC options
3.  a. assign river miles based on mainstem NHD
b. assign nearest river mile to things like RST, gages, based on lat/long
4.  visualize huc levels 
5.  apply to datasets to assign river miles, hucs 

```{r echo=TRUE, message=FALSE, warning=FALSE}
hucs <- read_sf(here::here('data-raw', 'shapefiles', 'WBDHU8_Klamath_Rogue.shp')) |> 
  filter(!grepl('17', huc8)) 

huc6 <- "180102"

ggplot() +
  geom_sf(data = hucs) +
  theme_minimal()

```

```{r echo=TRUE, message=FALSE, warning=FALSE}
layers <- st_layers(here::here('data-raw', 'shapefiles', 'NHDPLUS_H_1801_HU4_GDB.gdb'))

nhd_fcode <- st_read(here::here('data-raw', 'shapefiles', 'NHDPLUS_H_1801_HU4_GDB.gdb'), layer = 'NHDFCode') |> 
  janitor::clean_names() 

nhd_flowline_vaa<- st_read(here::here('data-raw', 'shapefiles', 'NHDPLUS_H_1801_HU4_GDB.gdb'), layer = 'NHDPlusFlowlineVAA') |> 
  janitor::clean_names() 

nhd_flowline_raw <- st_read(here::here('data-raw', 'shapefiles', 'NHDPLUS_H_1801_HU4_GDB.gdb'), layer = 'NHDFlowline') |> 
  janitor::clean_names() 

huc_12 <- st_read(here::here('data-raw', 'shapefiles', 'NHDPLUS_H_1801_HU4_GDB.gdb'), layer = 'WBDHU12') |> 
  janitor::clean_names() |> 
  st_transform(crs = "+proj=longlat +datum=WGS84") |> 
  filter(substr(huc12, 1, 6) == "180102")

huc_10 <- st_read(here::here('data-raw', 'shapefiles', 'NHDPLUS_H_1801_HU4_GDB.gdb'), layer = 'WBDHU10') |> 
  janitor::clean_names() |> 
  st_transform(crs = "+proj=longlat +datum=WGS84") |> 
  filter(substr(huc10, 1, 6) == "180102")

huc_8 <- st_read(here::here('data-raw', 'shapefiles', 'NHDPLUS_H_1801_HU4_GDB.gdb'), layer = 'WBDHU8') |> 
  janitor::clean_names() |> 
  st_transform(crs = "+proj=longlat +datum=WGS84") |> 
  filter(substr(huc8, 1, 6) == "180102")


fcode_descr <- c("Stream/River", "Artificial Path",
                 "Stream/River: Hydrographic Category = Intermittent",
                 "Stream/River: Hydrographic Category = Perennial", "Stream/River: Hydrographic Category = Ephemeral")  

nhd_flowline <- nhd_flowline_raw |> 
  mutate(rc_huc_6 = substr(reach_code, 1, 6),
         rc_huc_8 = substr(reach_code, 1, 8),
         rc_huc_10 = substr(reach_code, 1, 10),
         rc_huc_12 = substr(reach_code, 1, 12)) |> 
  filter(rc_huc_6 == "180102") |> # filter to klamath huc6
  filter(!is.na(gnis_name)) |> # remove NA streams 
  left_join(nhd_fcode) |> 
  filter(description %in% fcode_descr) |> 
  left_join(nhd_flowline_vaa) |> 
  glimpse()

ggplot() +
  geom_sf(data = hucs, aes(color = 'darkred')) +
  geom_sf(data = nhd_flowline) +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r message=FALSE, warning=FALSE}
# filtering to mainstem and tribs of interest 

rivers <- c('Klamath River', 'Trinity River', 'Sprague River',
            'Williamson River', 'Wood River', 'Lost River', 'Shasta River', 
            'Scott River', 'Salmon River', 'Blue Creek', 'Bogus Creek', 'Clear Creek',
            'Indian Creek', 'Link River')

nhd_flowline_filtered <- nhd_flowline |> 
  filter(gnis_name %in% rivers) |> 
  st_transform(crs = "+proj=longlat +datum=WGS84") |> 
  st_zm()

# TODO needs further cleaning of lines 

# ggplot() +
#   geom_sf(data = hucs) +
#   geom_sf(data = nhd_flowline_filtered) +
#   theme_minimal()

```

```{r}
leaflet(data = nhd_flowline_filtered) |> 
  addTiles() |> 
  addPolylines(popup = ~gnis_name)
```

### HUC exploration 

```{r}
leaflet() |> 
  addTiles() |> 
  addPolygons(data = huc_12,
              color = "darkgreen",
              popup = ~paste0("HUC12", "<br>", name),
              group = "HUC12") |> 
  addPolygons(data = huc_10,
              color = 'darkblue',
              popup = ~paste0("HUC10", "<br>", name),
              group = "HUC10") |> 
  addPolygons(data = huc_8,
              color = 'darkred',
              popup = ~paste0("HUC8", "<br>", name),
              group = "HUC8") |> 
  addPolylines(data = nhd_flowline_filtered,
               color = 'blue',
               popup = ~gnis_name,
               group = "Flowlines") |> 
  addLayersControl(
    overlayGroups = c("HUC12", "HUC10", "HUC8", "Flowlines"),
    options = layersControlOptions(collapsed = FALSE)
  )

```

### Assign River Miles 

```{r}
get_river_miles <- function(river_layer, interval_no = 10) {
  
  interval <- interval_no #interval_no is 10 meters by default
  line_length <- st_length(river_layer)
  n_points <- as.numeric(line_length %/% interval)
  
  river_points <- st_line_sample(river_layer, n = n_points)  |> 
    st_cast("POINT")
  
  river_points <- st_sf(geometry = river_points, crs = st_crs(river_layer))
  
  # units are in meters
  distances <- sapply(1:(length(river_points$geometry) - 1), function(i) {
    st_distance(river_points$geometry[i], river_points$geometry[i + 1])
  }) 
  cumulative_distances <- c(0, cumsum(distances)) |> sort(decreasing = TRUE)
  
  river_mile_points <- river_points  |> 
    mutate(river_meter = cumulative_distances,
           river_km = river_meter/1000,
           river_mile = river_meter/1609.34) |> 
    st_transform(crs = "+proj=longlat +datum=WGS84") 
  
  return(river_mile_points)
  
}

```

```{r}
# pre-processing: 
river <- nhd_flowline_filtered |> 
  filter(gnis_name == "Bogus Creek") |> 
  st_combine() |> 
  st_line_merge() |> 
  st_transform(crs = 32610)

river_mile_points <- get_river_miles(river, interval_no = 10)

leaflet() |> 
  addTiles() |> 
  addCircleMarkers(data = river_mile_points,
                   radius = ~river_mile / max(river_mile) * 10,
                   popup = ~as.character(round(river_mile, 2))) 


```

### Identify nearest river mile by lat/long   

```{r} 
get_nearest_river_mile <- function(river_mile_points, random_point) {
  distances <- st_distance(random_point, river_mile_points)
  min_distance_index <- which.min(distances)
  
  nearest_river_mile <- river_mile_points[min_distance_index, ]
  
  return(nearest_river_mile$river_mile)
}

random_lat <- 41.928313
random_long <- -122.430081
random_point <- st_sfc(st_point(c(random_long, random_lat)), crs = 4326)

get_nearest_river_mile(river_mile_points, random_point)

leaflet() |> 
  addTiles() |> 
  addCircleMarkers(data = river_mile_points,
                   radius = ~river_mile / max(river_mile) * 10,
                   popup = ~as.character(round(river_mile, 2))) |> 
  addMarkers(data = random_point)

```

```{r}
knitr::knit_exit()
```

### Convert to line layer

```{r}
river_mile_line <- river_mile_points |> 
  group_by(river_mile) |> 
  summarise(
    river_mile = first(river_mile),  # Retain the first value of an attribute
    do_union = FALSE
  ) |> 
  st_cast("LINESTRING")


ggplot() +
  geom_sf(data = river_mile_line)

leaflet() |> 
  addTiles() |> 
  addPolylines(data = river_mile_line) 

plot(river_mile_line)

```



```{r}
# combine the NHD layer with the points

# Identify the appropriate UTM zone for your data
utm_crs <- 32610 # Replace with the correct EPSG code for your area

# Transform both layers to the same CRS
river_mile_points <- st_transform(river_mile_points, utm_crs)
bogus_crk <- st_transform(bogus_crk, utm_crs)


snapped_points <- st_snap(river_mile_points, bogus_crk, tolerance = 10) # Adjust tolerance as needed

snapped_points <- st_cast(snapped_points, "POINT") # Ensure points are POINT geometries

# Optionally, sort points by an attribute or spatially
snapped_points <- snapped_points[order(snapped_points$river_mile), ] # Replace 'river_mile' with your column

snapped_line <- st_cast(st_combine(snapped_points), "LINESTRING")

st_crs(snapped_line) <- st_crs(bogus_crk)

ggplot() +
  geom_sf(data = snapped_line, alpha = 0.5) +
  geom_sf(data = bogus_crk) + 
  theme_minimal()

```

```{r}
test <- st_union(bogus_crk, river_mile_points) |> 
  #filter(river_mile > 0) |> 
  select(permanent_identifier, river_mile, hydro_seq, rc_huc_6,rc_huc_8, rc_huc_10, rc_huc_12) 

more_test <- st_collection_extract(test, type = c("LINESTRING"))


ggplot() +
  geom_sf(data = more_test, aes(linewidth = river_mile))

```

```{r eval=FALSE, include=FALSE}
# attempting to assign river miles by segment 
# 
# Filter for Bogus Creek
bogus_crk <- nhd_flowline_filtered |> 
  filter(gnis_name == "Bogus Creek") |> 
  arrange(hydro_seq) |>  # Arrange by hydro_seq for logical order
  mutate(segment_id = row_number())

# check to make sure the segments are going in the right direction 
ggplot() +
  geom_sf(data = bogus_crk, aes(linewidth = segment_id)) +
  theme_minimal()

# Transform to UTM for distance calculations
bogus_crk <- st_transform(bogus_crk, crs = 32610)

calc_distances <- function(id, river_layer) {
  
  segment <- river_layer |> 
    filter(segment_id == id) |> 
    st_cast("LINESTRING")
  
  # Set the sampling interval (250 meters in this example)
  interval <- 50
  
  # Calculate the total length of the line
  line_length <- st_length(segment)
  
  # Determine the number of points needed
  n_points <- as.numeric(line_length %/% interval)
  
  if(n_points == 0) {
    n_points = 2
  }
  
  # Sample points along the segment
  river_points <- st_line_sample(segment, n = n_points) |> #4 per km 
    st_cast("POINT")
  
  # Convert to sf object
  river_points <- st_sf(geometry = river_points, crs = st_crs(segment))
  
  ggplot() + 
    geom_sf(data = river_points) +
    geom_sf(data = segment)
  
  # Calculate distances between consecutive points
  distances <- sapply(1:(length(river_points$geometry) - 1), function(i) {
    st_distance(river_points$geometry[i], river_points$geometry[i + 1])
  })
  
  return(distances)
  
}

# Function to calculate river mile points for each segment
calculate_river_miles <- function(id, river_layer) {
  
  prev_distances <- calc_distances(id = id-1, river_layer = bogus_crk)
  distances <- calc_distances(id, river_layer = bogus_crk)
  
  # Calculate cumulative distances
  if(segment$segment_id == 1) {
    cumulative_distances <- c(0, cumsum(distances))
  } else {
    prev_segment <- segment_id - 1
  }
  
  # Add attributes to the points
  river_points <- river_points |> 
    mutate(
      river_meter = cumulative_distances,
      river_km = river_meter / 1000,
      river_mile = river_meter / 1609.34
    )
  
  
  return(river_points)
}

river_mile_points_list <- lapply(1:nrow(bogus_crk), function(i) {
  calculate_river_miles(bogus_crk[i, ])
})

# Combine the results
river_mile_points <- do.call(rbind, river_mile_points_list)

bogus_crk <- st_transform(bogus_crk, st_crs(river_mile_points)) |> 
  st_cast("POINT") |> 
  mutate(vertex_id = row_number()) 

```


```{r}

test_all <- bogus_crk |> 
  st_cast("LINESTRING") |> 
  st_line_sample(n = 1000)

river_points <- st_sf(geometry = test_all, crs = st_crs(bogus_crk))

ggplot() + 
  geom_sf(data = river_points) +
  geom_sf(data = bogus_crk)

```



cleaning data - merge with NHDFCode - remove NAs from gnis_name - filter by mainpath and innetwork == 1 - filter by fcodes that are not rivers (note: KEEP artificial paths as they are river segments) - add 8,10,12 digit HUCs using script:
mutate(rc_huc_8 = substr(reachcode, 1, 8), rc_huc_10 = substr(reachcode, 1, 10), rc_huc_12 = substr(reachcode, 1, 12))

river miles - make a point layer for river miles (fractions of a river mile) - sf::st_line_sample() makes points along the line, assign some X feet spacing - order segments from downstream to upstream, cumulative length moving down segments, what is the river mile at start? - each trib gets its own sequence of river miles - sf::st_nearest_feature() to find the nearest river mile to each feature (such as RST)

value added attributes \> hydrosequence (hydroseq) - this is flow order, so arranging table by this should be upstream to downstream (or vice versa)

### Downloading NHD using nhdPlusTools

Using `nhdplusTools`. I found this to be much slower. You end up receiving many geodatabases by HUC4s that would need to be merged. The download process takes significantly longer than how it is outlined above. 
```{r eval=FALSE, include=FALSE}
#install.packages('nhdplusTools')
library(nhdplusTools)

download_nhdplushr(here::here('data-raw', 'shapefiles'), c('180102'), download_files = TRUE)

```

```{r eval=FALSE, include=FALSE}
layers <- st_layers('data-raw/shapefiles/18/NHDPLUS_H_1810_HU4_GDB.gdb/')

data <- st_read('data-raw/shapefiles/18/NHD_H_1801_HU4_GDB.gdb/', layer = 'NHDFlowline')

ggplot() +
  geom_sf(data = data) +
  geom_sf(data = hucs, aes(color = 'red')) 

```


```{r}
#hucs
hucs <- sf::read_sf('shiny/data/shapefiles/WBDHU8_Klamath_Rogue_prj.shp') |> 
  sf::st_transform('+proj=longlat +datum=WGS84')
leaflet(data = hucs) |> 
  addTiles() |> 
  addPolygons(popup = ~name)

```
