---
title: "Klamath [Data type] [QC]"
author: "[Your name]"
date: "[Date]"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)
```

# Klamath [Data type long version]

## Description of Monitoring Data

**Timeframe:** 

**Completeness of Record throughout timeframe:** 

**Sampling Location:**

**Data Contact:** 

Any additional info?

## Access Data

```{r, eval=FALSE}
# read raw data file

```

Read in data from source, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
```

## Data transformations

```{r}
# For different excel sheets for each year read in and combine years here
```

```{r}
# Snake case, 
# Columns are appropriate types
# Remove redundant columns
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
```

### Variable: `[name]`

**Plotting [Variable] over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
```

**Numeric Summary of [Variable] over Period of Record**

```{r}
# Table with summary statistics
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Explore Categorical variables: {.tabset}

General notes: If there is an opportunity to turn yes no into boolean do so, but not if you loose value 

```{r}
# Filter clean data to show only categorical variables
```


### Variable: `[name]`
```{r}
#table() 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
```

**Create lookup rda for [variable] encoding:** 
```{r}
# Create named lookup vector
# Name rda [watershed]_[data type]_[variable_name].rda
# save rda to data/ 
```

**NA and Unknown Values**

Provide a stat on NA or unknown values

## Summary of identified issues

* List things that are funky but that we don't feel like should be changed without more investigation

## Save cleaned data back to /data folder 

```{r}
# Write to /data folder
# Name file [watershed]_[data type].csv
```