---
title: "Fall Chinook Escapement - Klamath Basin Watershed"
author: "[Your name]"
date: "[Date]"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(scales)
library(readxl)
```

The purpose of this markdown is to give an overview of the types of data available, temporal coverage, geographic coverage, and to get an idea of the amount of work that it would take us to clean it up and use it. 

This data was shared by [Morgan Knechtle](mailto::Morgan.Knechtle@wildlife.ca.gov) and it is publicly available at [California Open Data Portal](https://data.ca.gov/dataset/fall-chinook-escapement-klamath-basin-watershed) 

### Overview

Adult salmonid data collected from various areas within the Klamath basin watershed. Various tributaries are monitored using video systems (Bogus Creek, Shasta River, Scott River), other areas are monitored using redd/carcass surveys, and Iron gate Hatchery adult returns. The main purpose of the data is to estimate escapement of fall Chinook slamon, although data on other species (steelhead, coho salmon) also collected.

### Data Available {.tabset} 


#### Bogus Creek carcass spawning 

 - Title: Bogus Creek carcass spawning 
 - Extent: Survey is located on Bogus Creek from the confluence with the Klamath River upstream 3.6 miles.
 - Description: Data includes daily carcass recoveries from spawning grounds surveys. Biological data from carcasses includes date recovered, species, condition, reach collected, observed clips, fork length and biological sample collected.  Date utilized to describe population estimated at adult fish counting facility.
 - Initial collection: October 2008
 - Final collection: December 2022
 - Fields included: 
 
```{r, echo=FALSE}
bogus_carcass <- read_xlsx("data-raw/adult-escapement-cdfw/DMP_Bogus Creek Carcass Data 2008_2022.xlsx", sheet = 1)

str(bogus_carcass)
```

 - Name of xlsx file: DMP_Bogus Creek Carcass Data 2008_2022.xlsx
 
Cleaning implications:

 - Data in the xlsx is distributed by year per tab 
 - 2013 - 2022 data has the same fields, 2008 - 2012 are missing "condition" column. There are some minimal inconsistencies with column names accross years  
- Fields definition list:
  - Location and year
  - Date carcass encountered
  - Reach number
  - Species observed
  - Fork length measured in centimeters
  - Sex of fish observed
  - Determination of sucessful spawning
  - description of hatchery marks
  - unique number associated with head
  - was carcass included in systematic sample
  - Condition description
  - Unique sample number of each carcass
  - Where scales collected
  - Was tissue collected
  - Where otoliths collected
  - unique number for the record
  - Systematic sample rate
 
 
#### Upper Salmon River Carcass Mark Recapture Spawning
 
 - Title: Upper Salmon River Carcass Mark Recapture Spawning Grounds Surveys 
 - Extent: Survey is located on the Upper Salmon River (above Nordheimer Campground), including Upper Mainstem, and North and South Forks of the Salmon River and tributaries.
 - Description: Data includes daily carcass recoveries from spawning grounds surveys. Biological data from carcasses, disposition, and jaw tagging information used in a mark recapture population estimation study. Data provided used in Program R Cormack Jolly Seber popoulation estimator. Contemporary data set.
 - Initial collection: October 2011
 - Final collection: December 2022
 - Fields included:
  
```{r, echo=FALSE}
salmon_carcass <- read_xlsx("data-raw/adult-escapement-cdfw/Salmon River Carcass Mark Recapture 2011-2022.xlsx", skip = 1, sheet = 1)

str(salmon_carcass)
```

- Name of xlsx file: Salmon River Carcass Mark Recapture 2011-2022.xlsx
 
Cleaning implications:

 - Data in the xlsx is distributed by year per tab
 - Column names accross the different years are pretty similar
 
 - Fields definition list:
 
    - Date carcass encountered
    - Numerical value dispositon of carcass
    - Fork length measured in centimeters
    - Sex of fish observed
    - Determination of sucessful spawning
    - unique tag # attached to carcass upon first encounter
    - unique tag # of tag on carcass during recapture
    - release or re-released tagged fish, or chopped untagged fish
    - description of hatchery marks
    - unique identified tag applied upon recovery of known hatchery origin fish

