---
title: "Upper Klamath Lake sucker raw capture data"
author: "Ashley Vizek"
date: "February 2025"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(scales)
library(readxl)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", "#D67236",# Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87", # Isle of dogs 2 altered slightly
                  "#F3DF6C", "#CEAB07", "#D5D5D3", "#24281A", # Moonriese 1, 
                  "#798E87", "#C27D38", "#CCC591", "#29211F", # moonrise 2
                  "#85D4E3", "#F4B5BD", "#9C964A", "#CDC08C", "#FAD77B" ) # moonrise 3 

colors_two <- c("#D8A499", "#74A089")

# define AWS data bucket
# note that you need to set up access keys in R environ
klamath_project_board <- pins::board_s3(
  bucket="klamath-sdm",
  access_key=Sys.getenv("aws_access_key_id"),
  secret_access_key=Sys.getenv("secret_access_key_id"),
  session_token = Sys.getenv("session_token_id"),
  region = "us-east-1"
)
```

The purpose of this markdown is to give an overview of the dataset including specific data types, temporal coverage, and geographic coverage. This document is helpful in developing schemas to bring together disparate datasets and exploring data limitations.

# Overview

**Data type:** 

These data were manually downloaded from USGS ScienceBase Catalog: 

[Krause, J.R., and Paul-Wilson, R.K., 2024. Status and trends of adult Lost River (Deltistes luxatus) and shortnose (Chasmistes brevirostris) sucker populations in Upper Klamath Lake, Oregon, 2024 (ver. 2.0, October 2024). U.S. Geological Survey data release, https://doi.org/10.5066/P14AZJX5.](https://www.sciencebase.gov/catalog/item/670e6216d34edd269209562d)

We acquired these data as a preliminary exploration of sucker data and because they are an important long-term time series. 

USGS uses the Science Base catalog to publish datasets. There is an API and R package available however authentication is needed. I tested the `sbtools` package and downloading items using `item_file_download`. Some items downloaded but the data file didn't, perhaps because it was too large. In order to create an account you need to be sponsored by a USGS employee. Ideally we would use an API to acquire these datasets but for now they have been downloaded manually.

Helpful links:

- [ScienceBase documentation](https://www.usgs.gov/sciencebase-instructions-and-documentation/api-and-web-services)

## Data Available {.tabset} 

**Adult recaptures**

**Releases?**

Ideally we would also have releases or number tagged

### Sucker adult recaptures

```{r, include=FALSE}
recaptures_raw <- read_csv(here::here("data-raw", "fisheries", "USGS_Adult_Captures.csv"))
```

 - Title: Adult sucker captures
 - Spatial extent: Upper Klamath Lake
 - Description: Data were collected as part of a long-term capture-recapture program to assess the population dynamics of two long-lived, federally endangered catostomids in Upper Klamath Lake, Oregon. Lost River suckers (LRS; Deltistes luxatus) and shortnose suckers (SNS; Chasmistes brevirostris) have been captured and tagged with passive integrated transponder (PIT) tags during their spawning migrations in each year since 1995. Cormack-Jolly-Seber (CJS) open population capture-recapture models were used to estimate annual survival probabilities, and a reverse-time analog of the CJS model was used to estimate recruitment of new individuals into the spawning population. In addition, data on the size composition of captured fish were examined to provide corroborating evidence of recruitment. Separate analyses were done for each species and also for each subpopulation of LRS (i.e., lake and river spawners). Model estimates of survival and recruitment were used to derive estimates of changes in population size over time and to determine the status of the populations through 2024. Model estimates of survival, recruitment, and size composition are provided along with accompanying R code and input data. First release: 2023 Revised: October 2024 (ver. 2.0)
 - Temporal extent and resolution: 1995-2024, annual (ongoing?)
 - Fields included: `r colnames(recaptures_raw)`
 - Name of data file: ukl_adult_sucker_captures.csv
 
**Data processing implications:**

**Questions:**

- 

**Action items:**

- 

#### Data summary

This section may (or may not) explore every variable in the dataset. It will include high level summaries of the data to communicate the primary story of the data.

```{r, echo = F}
glimpse(recaptures_raw)
```

```{r, echo = FALSE}
recaptures_raw |> 
  mutate(year = year(datetime)) |> 
  group_by(year, length, prevailingtag) |> 
  tally() |> 
  ungroup() |> 
  group_by(year, length) |> 
  tally() |> 
  ggplot(aes(x = length, y = n)) +
  geom_col() +
  facet_wrap(~year)
```

### Scott and Shasta Coho juvenile abundance

```{r, include=FALSE}
juvenile_raw <- read_csv(here::here("data-raw", "fisheries", "modeled", "juvenile.csv"))
```


 - Title: Annual abundance estimates for Scott and Shasta Coho juveniles
 - Spatial extent: Scott and Shasta rivers
 - Description: Surveys of age-0+ (parr) and age-1+ (smolts) Coho Salmon emigrating from the Scott River have been conducted by CDFW since 2001 (Chesney and Knechtle 2016; Knechtle and Chesney 2016). Annual juvenile abundances were estimated from raw counts of fish captured in rotary screw traps located on the Scott and Shasta Rivers, 7.6 km and 0.1 km upstream of their respective confluences with the Klamath River. In most years, trap efficiencies were estimated from a mark-recapture framework and then used to estimate abundances from raw counts (Carlson et al. 1998). In years where inadequate numbers of Coho Salmon were captured or marked to estimate independent trap efficiencies, correlations between capture efficiencies of juvenile Coho Salmon and juvenile steelhead trout from prior years were instead used to estimate abundance. We estimated the freshwater productivity models with annual estimates of parr and smolt abundance in the Scott and Shasta Rivers, and we estimated the parr and smolt emigration timing models with weekly abundance estimates for each life stage.
 - Temporal extent and resolution: 2004-2013, annual
 - Fields included: `r colnames(juvenile_raw)`
 - Name of data file: juvenile.csv
 
**Data processing implications:**

**Questions:**

- What method is used to expand catch using trap efficiencies?
- What is meant by "correlations with steelhead trout"?
- How is catch expanded for days when the trap was not running?
- Are there uncertainty estimates around these numbers?

**Action items:**
- Try to find a more updated table with data through present
- Try to find if these data are available in other locations

#### Data summary

This section may (or may not) explore every variable in the dataset. It will include high level summaries of the data to communicate the primary story of the data.

```{r, echo = FALSE}
juvenile_raw |> 
  ggplot(aes(x = year, y = juvenile_abundance, colour = stream)) +
  geom_line() +
  scale_color_manual(values = colors_two) +
  theme_bw() +
  labs(x = "",
       y = "juvenile coho abundance estimate",
       color = "") +
  facet_wrap(~lifestage)

```

## Save data

Data files are stored on AWS

```{r, include = F}
klamath_project_board |> pins::pin_write(spawner_raw,
                               type = "csv",
                               title = "coho_adult_abundance")

klamath_project_board |> pins::pin_write(juvenile_raw,
                               type = "csv",
                               title = "coho_juvenile_abundance")
```