---
title: "Relevant Data USFWS Klamath Falls National Fish Hatchery Report - 2024"
author: "Badhia Yunes Katz (FlowWest)"
date: "`r Sys.Date()`"
output: html_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>", 
  fig.width=8, fig.height=5)
library(tidyverse)
library(lubridate)
library(tabulapdf)
library(janitor)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(plotly)
```

The purpose of this markdown is to highlight the data that we will use.

```{r include=FALSE}
pdf_path <- "data-raw/fisheries/hatcheries/20250619_FY2024 KFNFH Annual Report_Final Draft.pdf"
tables_raw <- extract_tables(pdf_path, method = "stream", output = "tibble")
```


```{r include=FALSE}
table_11_raw <- tables_raw[[13]] |>
  clean_names() |> 
  mutate(stock_date = tolower(stock_date)) |>
  glimpse()

table_11 <- table_11_raw |> 
  mutate(stock_date = as.character(stock_date),
         stock_date = na_if(stock_date, ""),
         date_anchor = if_else(!is.na(stock_date), stock_date, NA_character_))|> 
  fill(date_anchor, .direction = "down") |> 
  mutate(stock_date = if_else(is.na(stock_date), date_anchor, stock_date),
         stock_date = case_when(stock_date %in% c("overall", "distribution", "total") ~ "overall distrbution total",
                                TRUE ~ stock_date)) |> 
  select(-date_anchor)
```

```{r eval=FALSE, include=FALSE}
knitr::kable(table_11)
```

## Release totals 

```{r echo=FALSE}
# extracting the summaries/totals
release_totals <- table_11 |> 
  slice(19:24, 26, 29, 36:40, 42, 49:67) |> 
  select(-actual_tl_mm, -from_ponds, -stocking_location) |> 
  mutate(stock_date = stock_date |> 
      str_remove("\\s*totals?$") |> 
      str_squish()) 
```

```{r echo=FALSE}
knitr::kable(release_totals)
```


```{r echo=FALSE}
# Tol Muted color palette
tol_muted <- c("#332288", "#117733", "#44AA99", "#88CCEE",
               "#DDCC77", "#CC6677", "#AA4499", "#882255")

moth_release_totals <- release_totals |> 
  slice(1:21) |> 
  group_by(stock_date, species) |> 
  summarise(release_total = sum(number)) |> 
  ungroup()
```

Summarizing by month

```{r echo=FALSE}
plot_1 <- ggplot(moth_release_totals |> 
         slice(1:21),
       aes(x = stock_date, y = release_total, fill = species)) +
  geom_col() +
  # geom_text(aes(label = release_total),
  #           position = position_stack(vjust = 0.5),
  #           color = "black",
  #           size = 3) +
  labs(title = "Total Fish Released by Species",
       x = "Month of Release",
       y = "Number of Fish",
       fill = "Species") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = tol_muted)

ggplotly(plot_1)

```

```{r echo=FALSE}
knitr::kable(moth_release_totals)
```


```{r echo=FALSE}
release_totals_type <- release_totals |>
  slice(22:33) |> 
  mutate(release_type = case_when(
    str_detect(lot, regex("prod|production", ignore_case = TRUE)) ~ "Production",
    str_detect(lot, regex("fry", ignore_case = TRUE)) ~ "Fry",
    str_detect(lot, regex("salv", ignore_case = TRUE)) ~ "Salvage",
    TRUE ~ NA_character_)) |> 
  group_by(stock_date, species) |> 
  summarise(total_released = sum(number, na.rm = TRUE),
            .groups = "drop")
  
ggplot(release_totals_type,
       aes(x = species, y = total_released, fill = stock_date)) +
  geom_col(position = position_dodge(width = 0.8)) +
  geom_text(aes(label = scales::comma(total_released)),
            position = position_dodge(width = 0.8),
            vjust = -0.3,
            size = 3) +
  labs(title = "FY24 Species Totals vs Overall Distribution",
    x = "Species / Category",
    y = "Number of Fish Released",
    fill = "Group") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = tol_muted)


```


## Upper Klamath Lake sucker net pen survival
 
Overall Survival: 28.2% (Historically) and FY23 Survival: 2.9% and 3.1%

These very low historical survival rates (and the decision not to stock production fish in FY24) indicate the failure of the UKL net pen as a reliable growout and acclimation tool for production fish


```{r include=FALSE}
table_9_raw <- tables_raw[[11]] |> 
  clean_names() |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-2, ] |>
      setNames(new_names)
    })()
```

```{r echo=FALSE}
ukl_pen_survival <- table_9_raw |> 
  slice(2:12) |> 
  separate(stock_tl_number, into = c("stock_tl_mm", "number_stocked"), sep = "\\s+", remove = TRUE) |> 
  mutate(operation_year = as.numeric(operation),
         captured_year = as.numeric(x2_fish_cy),
         pen_id = x3_pen_id,
         stock_tl_mm = as.numeric(stock_tl_mm),
         number_stocked = as.numeric(number_stocked),
         harvest_tl_mm = as.numeric(harvest_tl),
         number_harvested = as.numeric(number),
         survival_percentage = as.numeric(x8_survival_percent)) |> 
  select(operation_year, captured_year, pen_id, stock_tl_mm, number_stocked, harvest_tl_mm,
         number_harvested, survival_percentage) 


```

```{r echo=FALSE}
ggplot(ukl_pen_survival, aes(x = factor(operation_year), y = survival_percentage, fill = pen_id)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  labs(x = "Operation Year",
    y = "Survival Percentage",
    fill = "Pen ID",
    title = "Survival Percentage by Pen and Year"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = tol_muted)
```

```{r echo=FALSE}
knitr::kable(ukl_pen_survival)
```

## Gerber net pen sucker survival and growth data

Successful external rearing and acclimation in Gerber Reservoir during the FY24 growout season. 96.3% and
 94.9%

Overall Survival: 83.9% (Historically, since 2021)

```{r include=FALSE}
table_10_raw <- tables_raw[[12]] |> 
  clean_names() |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-2, ] |>
      setNames(new_names)
    })()
```

Plotting both net pens

```{r include=FALSE}
gerber_pen <- gerber_pen_survival |> 
  mutate(location = "garber") |> 
  select(operation_year, captured_year, number_harvested, survival_percentage, location) 

ukl_pen <- ukl_pen_survival |> 
  mutate(location = "ukl") |> 
  select(operation_year, captured_year, number_harvested, survival_percentage, location) 

pen <- bind_rows(ukl_pen, gerber_pen) |> glimpse()


point_plot <- ggplot(pen, aes(x = operation_year, y = survival_percentage, color = location, 
                              text = paste0("Operation Year: ", operation_year, "<br>",
                                            "Captured Year: ", captured_year, "<br>",
                                            "Number Harvested: ", number_harvested, "<br>",
                                            "Survival Percentage: ", survival_percentage))) +
  geom_point() +
  theme_minimal() 


ggplotly(point_plot, tooltip = "text")

```

## Historical Collections and Releases Summary

```{r include=FALSE}
table_1_raw <- tables_raw[[2]] |>
  clean_names() |> 
  glimpse()

table_1 <- table_1_raw |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names) |>
      mutate(across(-1, ~parse_number(.x)))
    })()

historical_collections_releases <- table_1 |> 
  rename(sarp_tl_mm = tl_4_mm,
         fingerling_tl_mm = tl_6_mm,
         fry_tl_mm = tl_8_mm,
         salvage_sl_mm = sl_or_tl_mm) |> 
  slice(-10) |> # removing the total 
  mutate(fiscal_year = as.numeric(fiscal_year),
         salvage_tl_mm = case_when(salvage_sl_mm == 129 ~ salvage_sl_mm,
                                   salvage_sl_mm == 126 ~ salvage_sl_mm,
                                   TRUE ~ NA),
         salvage_sl_mm = case_when(salvage_tl_mm == 129 ~ NA,
                                   salvage_tl_mm == 126 ~ NA,
                                   TRUE ~ salvage_sl_mm))|> 
  glimpse()
```


```{r echo=FALSE}
knitr::kable(historical_collections_releases)
```


## Hatching Rate

```{r include=FALSE}
#table 3 - unsure on if we should include this table
table_3_raw <- tables_raw[[4]]
```

```{r echo=FALSE}
incubation_hatch_raw <- table_3_raw |> 
  clean_names() |> 
  mutate(spawning_date = as.Date(spawning_date, format = "%m/%d/%Y"),
         hatch_date = as.Date(hatch_date, format = "%m/%d/%Y")) |>
  # select(-x6) |> 
  # slice(-15) |> # removing the totals for now
  # rename(male_1 = male_3,
  #        male_2 = male_4,
  #        male_3 = male_male) |> 
  select(spawning_date, egg_volume_m_l, number_eggs_m_l, total_eggs, total_fry, hatch_percent, hatch_date) |> 
  glimpse()
```
