---
title: "Relevant Data USFWS Klamath Falls National Fish Hatchery Report - 2024"
author: "Badhia Yunes Katz (FlowWest)"
date: "`r Sys.Date()`"
output: html_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>", 
  fig.width=8, fig.height=5)
library(tidyverse)
library(lubridate)
library(tabulapdf)
library(janitor)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(plotly)
```

The purpose of this markdown is to highlight the data that we will use from the 2024 Hatchery Report.

Species: Lost River Sucker (Deltistes luxatus, LRS), Shortnose Sucker (Chasmistes brevirostris, SNS) 
```{r include=FALSE}
pdf_path <- "data-raw/fisheries/hatcheries/20250619_FY2024 KFNFH Annual Report_Final Draft.pdf"
tables_raw <- extract_tables(pdf_path, method = "stream", output = "tibble")
```


```{r include=FALSE}
table_11_raw <- tables_raw[[13]] |>
  clean_names() |> 
  mutate(stock_date = tolower(stock_date)) |>
  glimpse()

table_11 <- table_11_raw |> 
  mutate(stock_date = as.character(stock_date),
         stock_date = na_if(stock_date, ""),
         date_anchor = if_else(!is.na(stock_date), stock_date, NA_character_))|> 
  fill(date_anchor, .direction = "down") |> 
  mutate(stock_date = if_else(is.na(stock_date), date_anchor, stock_date),
         stock_date = case_when(stock_date %in% c("overall", "distribution", "total") ~ "overall distrbution total",
                                TRUE ~ stock_date)) |> 
  select(-date_anchor)
```

```{r eval=FALSE, include=FALSE}
knitr::kable(table_11)
```

## Release totals (raw)

- Table below shows the monthly summaries, and total fiscal year summary. 
- There is finer resolution data about the actual stock date by lot, and which pond and stocking location fish came from. I think Monthly totals are probably what we want. I will remove the overall totals at the bottom of this table 

```{r include=FALSE}
# extracting the summaries/totals
release_totals <- table_11 |> 
  slice(19:24, 26, 29, 36:40, 42, 49:67) |> 
  select(-actual_tl_mm, -from_ponds, -stocking_location) |> 
  mutate(stock_date = stock_date |> 
      str_remove("\\s*totals?$") |> 
      str_squish()) |> 
  # mutate(stock_month = stock_date) |> 
  # select(stock_month, species, lot, number) |>
  # slice(1:21) |> 
  glimpse()
```

```{r echo=FALSE}
knitr::kable(release_totals)
```


```{r echo=FALSE}
# Tol Muted color palette
tol_muted <- c("#332288", "#117733", "#44AA99", "#88CCEE",
               "#DDCC77", "#CC6677", "#AA4499", "#882255")

moth_release_totals <- release_totals |> 
  slice(1:21) |> 
  mutate(species = tolower(species)) |> 
  group_by(stock_date, species) |> 
  summarise(release_total = sum(number)) |> 
  ungroup() |> 
  rename(stock_month = stock_date)
```

## Release monthly summary totals - Data to be used

```{r echo=FALSE}
plot_1 <- ggplot(moth_release_totals |> 
         slice(1:21),
       aes(x = stock_month, y = release_total, fill = species)) +
  geom_col() +
  # geom_text(aes(label = release_total),
  #           position = position_stack(vjust = 0.5),
  #           color = "black",
  #           size = 3) +
  labs(title = "Total Fish Released by Species",
       x = "Month of Release",
       y = "Number of Fish",
       fill = "Species") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = tol_muted)

ggplotly(plot_1)

```

Final shaped data table looks like this:

```{r echo=FALSE}
knitr::kable(moth_release_totals)
```


```{r include=FALSE}
# save data
# usethis::use_data(moth_release_totals, overwrite = TRUE)
```


```{r eval=FALSE, include=FALSE}
# not sure the overall totals make sense to keep

release_totals_type <- release_totals |>
  slice(22:33) |> 
  mutate(release_type = case_when(
    str_detect(lot, regex("prod|production", ignore_case = TRUE)) ~ "Production",
    str_detect(lot, regex("fry", ignore_case = TRUE)) ~ "Fry",
    str_detect(lot, regex("salv", ignore_case = TRUE)) ~ "Salvage",
    TRUE ~ NA_character_)) |> 
  group_by(stock_date, species) |> 
  summarise(total_released = sum(number, na.rm = TRUE),
            .groups = "drop")
  
ggplot(release_totals_type,
       aes(x = species, y = total_released, fill = stock_date)) +
  geom_col(position = position_dodge(width = 0.8)) +
  geom_text(aes(label = scales::comma(total_released)),
            position = position_dodge(width = 0.8),
            vjust = -0.3,
            size = 3) +
  labs(title = "FY24 Species Totals vs Overall Distribution",
    x = "Species / Category",
    y = "Number of Fish Released",
    fill = "Group") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = tol_muted)


```

## Historical Collections and Releases Summary

This data table will be helpful to understand the amount of larvae collected per fiscal year, and the lifestage of fish released. The standard lengths (SL) and total lengths (TL) are available for each release btu I am leaving them out 

```{r include=FALSE}
table_1_raw <- tables_raw[[2]] |>
  clean_names() |> 
  glimpse()

table_1 <- table_1_raw |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names) |>
      mutate(across(-1, ~parse_number(.x)))
    })()

historical_collections_releases_raw <- table_1 |> 
  rename(sarp_tl_mm = tl_4_mm,
         fingerling_tl_mm = tl_6_mm,
         fry_tl_mm = tl_8_mm,
         salvage_sl_mm = sl_or_tl_mm) |> 
  slice(-10) |> # removing the total 
  mutate(fiscal_year = as.numeric(fiscal_year),
         salvage_tl_mm = case_when(salvage_sl_mm == 129 ~ salvage_sl_mm,
                                   salvage_sl_mm == 126 ~ salvage_sl_mm,
                                   TRUE ~ NA),
         salvage_sl_mm = case_when(salvage_tl_mm == 129 ~ NA,
                                   salvage_tl_mm == 126 ~ NA,
                                   TRUE ~ salvage_sl_mm))|> 
  glimpse()
```

```{r include=FALSE}
historical_collections_releases <- historical_collections_releases_raw |> 
  select(-c(sarp_tl_mm, fingerling_tl_mm, fry_tl_mm, salvage_sl_mm, salvage_tl_mm)) |> 
  glimpse()
```

Larvae Collected vs. Release Types by Fiscal Year

```{r echo=FALSE}
# wide to long format
releases_long <- historical_collections_releases |>
  pivot_longer(cols = -fiscal_year,
               names_to = "release_type",
               values_to = "count")

plot_2 <- ggplot(releases_long, aes(x = factor(fiscal_year), y = count, fill = release_type)) +
  geom_col(position = "dodge") +
  scale_fill_brewer(palette = "Set2") +
  labs(x = "Fiscal Year",
       y = "Count",
       fill = "Release Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 14, face = "bold")) +
  scale_fill_manual(values = tol_muted)

ggplotly(plot_2)
```

```{r eval=FALSE, include=FALSE}
# option to shape table
# historical_collections_releases_long <- historical_collections_releases |> 
#   pivot_longer(cols = c(sarp_release, fingerling_release, fry_release, salvage_release),
#                names_to = "lifestage",
#                values_to = "release") |> 
#   mutate(lifestage_released = case_when(lifestage == "sarp_release" ~ "sarp",
#                                lifestage == "fingerling_release" ~ "fingerling",
#                                lifestage == "fry_release" ~ "fry",
#                                lifestage == "salvage_release" ~ "salvage")) |> 
#   pivot_longer(cols = c(sarp_tl_mm, fingerling_tl_mm, fry_tl_mm, salvage_tl_mm),
#                names_to = "tl_stage",
#                values_to = "tl_mm") |> 
#   mutate(tl_stage = case_when(tl_stage == "sarp_tl_mm" ~ "sarp",
#                               tl_stage == "fingerling_tl_mm" ~ "fingerling",
#                               tl_stage == "fry_tl_mm" ~ "fry",
#                               tl_stage == "salvage_tl_mm" ~ "salvage")) |> 
#   filter(lifestage_released == tl_stage) |> 
#   select(fiscal_year, larvae_collected, lifestage_released, tl_mm) |> 
#   glimpse()
```


```{r echo=FALSE}
knitr::kable(historical_collections_releases)
```

```{r include=FALSE}
# save data
# usethis::use_data(historical_collections_releases, overwrite = TRUE)
```


## Upper Klamath Lake sucker net pen survival (unsure we need)
 
Overall Survival: 28.2% (Historically) and FY23 Survival: 2.9% and 3.1%

These very low historical survival rates (and the decision not to stock production fish in FY24) indicate the failure of the UKL net pen as a reliable growout and acclimation tool for production fish


```{r include=FALSE}
table_9_raw <- tables_raw[[11]] |> 
  clean_names() |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-2, ] |>
      setNames(new_names)
    })()
```

```{r echo=FALSE}
ukl_pen_survival <- table_9_raw |> 
  slice(2:12) |> 
  separate(stock_tl_number, into = c("stock_tl_mm", "number_stocked"), sep = "\\s+", remove = TRUE) |> 
  mutate(operation_year = as.numeric(operation),
         captured_year = as.numeric(x2_fish_cy),
         pen_id = x3_pen_id,
         stock_tl_mm = as.numeric(stock_tl_mm),
         number_stocked = as.numeric(number_stocked),
         harvest_tl_mm = as.numeric(harvest_tl),
         number_harvested = as.numeric(number),
         survival_percentage = as.numeric(x8_survival_percent)) |> 
  select(operation_year, captured_year, pen_id, stock_tl_mm, number_stocked, harvest_tl_mm,
         number_harvested, survival_percentage) 


```

```{r echo=FALSE}
ggplot(ukl_pen_survival, aes(x = factor(operation_year), y = survival_percentage, fill = pen_id)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  labs(x = "Operation Year",
    y = "Survival Percentage",
    fill = "Pen ID",
    title = "Survival Percentage by Pen and Year"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = tol_muted)
```

```{r echo=FALSE}
knitr::kable(ukl_pen_survival)
```

## Gerber net pen sucker survival and growth data (unsure we need)

Successful external rearing and acclimation in Gerber Reservoir during the FY24 growout season. 96.3% and
 94.9%

Overall Survival: 83.9% (Historically, since 2021)

```{r include=FALSE}
table_10_raw <- tables_raw[[12]] |> 
  clean_names() |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-2, ] |>
      setNames(new_names)
    })()

gerber_pen_survival <- table_10_raw |> 
  slice(2:10) |> 
  mutate(operation_year = as.numeric(operation),
         captured_year = as.numeric(fish),
         pen_id = x3_pen_id,
         stock_tl_mm = as.numeric(stock_tl),
         number_stocked = as.numeric(x5_number_stocked),
         harvest_tl_mm = as.numeric(harvest_tl),
         number_harvested = as.numeric(x7_number_harvested),
         survival_percentage = as.numeric(x8_survival_percent)) |> 
  select(operation_year, captured_year, pen_id, stock_tl_mm, number_stocked, harvest_tl_mm,
         number_harvested, survival_percentage) 
```
```{r echo=FALSE}
ggplot(gerber_pen_survival, aes(x = factor(operation_year), y = survival_percentage, fill = pen_id)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  labs(x = "Operation Year",
    y = "Survival Percentage",
    fill = "Pen ID",
    title = "Survival Percentage by Pen and Year"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = tol_muted)
```

Plotting both net pens (UKL and Gerber)

```{r include=FALSE}
gerber_pen <- gerber_pen_survival |> 
  mutate(location = "garber") |> 
  select(operation_year, captured_year, number_harvested, survival_percentage, location) 

ukl_pen <- ukl_pen_survival |> 
  mutate(location = "ukl") |> 
  select(operation_year, captured_year, number_harvested, survival_percentage, location) 

pen <- bind_rows(ukl_pen, gerber_pen) |> glimpse()


point_plot <- ggplot(pen, aes(x = operation_year, y = survival_percentage, color = location, 
                              text = paste0("Operation Year: ", operation_year, "<br>",
                                            "Captured Year: ", captured_year, "<br>",
                                            "Number Harvested: ", number_harvested, "<br>",
                                            "Survival Percentage: ", survival_percentage))) +
  geom_point() +
  theme_minimal() +
   scale_color_manual(values = tol_muted)
  
```


```{r echo=FALSE}
ggplotly(point_plot, tooltip = "text")
```



```{r include=FALSE}
## Hatching Rate
#table 3 - unsure on if we should include this table
table_3_raw <- tables_raw[[4]]
```

```{r eval=FALSE, include=FALSE}
incubation_hatch_raw <- table_3_raw |> 
  clean_names() |> 
  mutate(spawning_date = as.Date(spawning_date, format = "%m/%d/%Y"),
         hatch_date = as.Date(hatch_date, format = "%m/%d/%Y")) |>
  # select(-x6) |> 
  # slice(-15) |> # removing the totals for now
  # rename(male_1 = male_3,
  #        male_2 = male_4,
  #        male_3 = male_male) |> 
  select(spawning_date, egg_volume_m_l, number_eggs_m_l, total_eggs, total_fry, hatch_percent, hatch_date) |> 
  glimpse()
```
