---
title: "USFWS Klamath Falls National Fish Hatchery Reports"
author: "Badhia Yunes Katz (FlowWest)"
date: "`r Sys.Date()`"
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>", 
  fig.width=8, fig.height=5)
library(tidyverse)
library(lubridate)
library(tabulapdf)
library(janitor)
library(dplyr)
library(tidyr)
library(readr)
```

The tables in the Klamath Falls National Fish Hatchery (KFNFH) annual reports are designed to provide detailed documentation of operations and results. These reports were prepared by Prepared by the Klamath Falls National Fish Hatchery.

This markdown will extract and explore the tables on the hatchery table reports. 

It will identify any questions/clarification we might need as we go through the reports.

**Link to Sources**

## Annual Report for Fiscal Year 2021


### Table 1

Collection and release information for the Sucker Assisted Rearing Program (SARP) since its inception in 2016 by fiscal years. This table tracks key metrics over time, including Larvae Collected, SARP Release numbers, and Salvage Release numbers.

```{r echo=FALSE}
# extract tables
pdf_path_2021 <- "data-raw/fisheries/hatcheries/20240912_FY2021_KFNFH_Annual_Report_Final.pdf"
tables_2021 <- extract_tables(pdf_path_2021, method = "stream", output = "tibble")

table_1 <- tables_2021[[1]] |>
  clean_names()

# fixing so col name grabs the first row too
table_2021_clean <- tables_2021[[1]] |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names) |>
      mutate(across(-1, ~parse_number(.x)))
    })()

knitr::kable(table_2021_clean)

```

### Table 2

*Summary of Lost River Sucker (LRS), Shortnose Sucker (SNS), and Klamath Largescale Sucker (KLS) stocked during the Fiscal Year 2020 (October 1, 2020, to September 30, 2021).* This table breaks down fish stocking activities by date, species, Lot #, weight, density (#/lb.), Total Length (TL) in inches and mm, the ponds the fish originated from, and the Stocking Location.

```{r echo=FALSE}
table_2 <- tables_2021[[2]] |>
  clean_names() |> 
  mutate(is_real_date = grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", date), # fill with date
         date_lower = tolower(date), # detect full month names 
         is_month = date_lower %in% tolower(month.name),
         date_anchor = case_when(is_real_date ~ date,
                                 is_month ~ date, 
                                 TRUE ~ NA_character_)) |> 
  fill(date_anchor, .direction = "down") |> 
  mutate(date = if_else(is.na(date), date_anchor, date)) |> 
  select(-is_real_date, -is_month, -date_lower, -date_anchor) 

knitr::kable(table_2)
```

### Table 3

*Fall 2020 production summary.* This table summarizes activities during the fall of 2020, tracking fish cohorts (Lot #) by Harvest Date and Pond, noting inventory numbers (Paper # and Actual #), survival rates, the number and percentage of fish stocked, the number and percentage of fish graded as "Short," the number of fish designated as broodstock (# Brood), and the disposition of the fish

```{r echo=FALSE}
table_3 <- tables_2021[[3]] |>
  clean_names() |> 
  glimpse()
```

### Table 4

Spring 2021 production summary. 

```{r echo=FALSE}
table_4 <- tables_2021[[4]] |>
  clean_names() 
```

### Table 5 

Summary of PIT Tag Study Pond stockings in Replicate Block 1 during week of March 22, 2021.

```{r echo=FALSE}
table_5 <- tables_2021[[5]] |>
  clean_names() |>
    (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names) |>
      mutate(across(-1, ~parse_number(.x)))
    })()

knitr::kable(table_5)
```

### Table 6 

Summary of PIT Tag Study Pond stockings in Replicate Block 2 during week of March 29, 2021. 

```{r echo=FALSE}
table_6 <- tables_2021[[6]] |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    
    x[-1, ] |>
      setNames(new_names) |>
      mutate(
        across(
          # apply parse_number to all except first column and tagged_vs_untagged
          -c(1, tagged_vs_untagged),
          ~parse_number(.x)
        )
      )
  })()

knitr::kable(table_6)
```

### Table 7

Summary of PIT Tag Study Pond stockings in Replicate Block 3 during week of April 5, 2021.

```{r echo=FALSE}
table_7 <- tables_2021[[7]] |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    
    x[-1, ] |>
      setNames(new_names) |>
      mutate(
        across(
          # apply parse_number to all except first column and tagged_vs_untagged
          -c(1, tagged_vs_untagged),
          ~parse_number(.x)
        )
      )
  })()

knitr::kable(table_7)
```