---
title: "Most Recent USFWS Klamath Falls National Fish Hatchery Report - 2024"
author: "Badhia Yunes Katz (FlowWest)"
date: "`r Sys.Date()`"
output: html_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>", 
  fig.width=8, fig.height=5)
library(tidyverse)
library(lubridate)
library(tabulapdf)
library(janitor)
library(dplyr)
library(tidyr)
library(readr)
```

The purpose of this markdown is to understand the data available in the most recent Hatchery report, identify questions or data gaps. 

```{r include=FALSE}
pdf_path <- "data-raw/fisheries/hatcheries/20250619_FY2024 KFNFH Annual Report_Final Draft.pdf"
tables_raw <- extract_tables(pdf_path, method = "stream", output = "tibble")

table_1_raw <- tables_raw[[2]] |>
  clean_names() |> 
  glimpse()
```


## KFNFH Historical Collections and Releases Summary (2016–FY24)

 - Larvae collected: The total number of wild larval suckers collected, primarily from the lower Williamson River, for production purposes
 - SARP release: The number of "Production Fish" released. These are fish raised to the target size of approximately 200 mm total length (TL), typically taking 18–24 months
 - Fingerling release: The number of smaller, excess fish stocks (beyond rearing capacity) that were transferred or released into off-channel rearing locations
 - Fry release: The number of small fish released, generally produced from experimental or wild-spawning activities on the lake and river side, or transferred to other facilities
 - Salvage release: The number of suckers salvaged during the year (e.g., from Klamath Project Operations or dam removal efforts) that were rehabilitated, tagged, and then repatriated (released back to the wild)
 - TL vs SL: The average Standard Length (SL) or Total Length (TL) are both in millimeters for the salvaged fish, however, they are inconsistently collected- SL in earlier years (2018–2021) and TL in later years (2023–2024)

```{r include=FALSE}
table_1 <- table_1_raw |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names) |>
      mutate(across(-1, ~parse_number(.x)))
    })()

historical_collections_releases <- table_1 |> 
  rename(sarp_tl_mm = tl_4_mm,
         fingerling_tl_mm = tl_6_mm,
         fry_tl_mm = tl_8_mm,
         salvage_sl_mm = sl_or_tl_mm) |> 
  slice(-10) |> # removing the total 
  mutate(fiscal_year = as.numeric(fiscal_year),
         salvage_tl_mm = case_when(salvage_sl_mm == 129 ~ salvage_sl_mm,
                                   salvage_sl_mm == 126 ~ salvage_sl_mm,
                                   TRUE ~ NA),
         salvage_sl_mm = case_when(salvage_tl_mm == 129 ~ NA,
                                   salvage_tl_mm == 126 ~ NA,
                                   TRUE ~ salvage_sl_mm))|> 
  glimpse()
```


```{r echo=FALSE}
knitr::kable(historical_collections_releases)
```

## Wild Adult LRS Handled

Inventory of the individual wild adult Lost River Suckers (LRS) handled by KFNFH staff during the spawning collection efforts at the East Side Springs (ESS) of Upper Klamath Lake in 2024.
The table summarizes the wild LRS captured, identifying their physical characteristics, tagging information, and how their gametes were used.

 - Lists the four dates when gamete collection occurred: April 17, April 24, May 1, and May 8, 2024.
 - documents the total of 43 wild adult LRS handled (13 females and 30 males). All these fish were returned safely to the water thereafter.
 - confirms the sex of the fish and whether they were "Spawned" or "Not Spawned". Twenty-four of the 30 males were spawned.
  - gamete_use column notes they were "Cryo Preservation" or "Spawned/Cryopreservation," indicating that excess milt was used for cryopreservation trials.
  - family_groups: most females were crossed with four males, though one female was crossed with two, totaling 50 family groups collected overall. This column applies only to the females.
  
  
Success indicators:
-  Incubation Improvement: KFNFH staff achieved a significant milestone by increasing onsite hatching rates from around 20% to over 70%. This table shows information on the male and female Lost River Suckers used in the spawning effort and the incubation results for each female
- Successful captive LRS spawning: hatchery pawned and hatched an estimated 56,458 larval Lost River Suckers on station from gametes collected from the East Side Springs population. (this number is at the very bottom of the table but was not read in here. It is the sum of all rows under total_fry)

Failure indicators: 

```{r include=FALSE}
table_2_raw <- tables_raw[[3]]
```

```{r echo=FALSE}
wild_lrs_adults <- table_2_raw |> 
  clean_names() |> 
  mutate(date = as.Date(date, format = "%m/%d/%Y")) |> 
  glimpse()
```

## Summary of Incubation and Hatch Results

Quantitative results of the fertilization and incubation trials conducted using the gametes collected from the wild LRS documented in wild_lrs_adults table.

Success indicators:
-  Incubation Improvement: KFNFH staff achieved a significant milestone by increasing onsite hatching rates from around 20% to over 70%

Failure indicators: 

```{r include=FALSE}
#table 3
table_3_raw <- tables_raw[[4]]
```

```{r echo=FALSE}
incubation_hatch_raw <- table_3_raw |> 
  clean_names() |> 
  mutate(spawning_date = as.Date(spawning_date, format = "%m/%d/%Y"),
         hatch_date = as.Date(hatch_date, format = "%m/%d/%Y")) |>
  select(-x6) |> 
  slice(-15) |> # removing the totals for now
  rename(male_1 = male_3,
         male_2 = male_4,
         male_3 = male_male) |> 
  glimpse()
```

```{r echo=FALSE}
incubation_hatch_raw |>
  ggplot(aes(hatch_percent)) +
  geom_histogram() +
  theme_minimal()
```

Mean Hatch (%) grouped by the four Spawning Dates (April 17, 24, May 1, May 8)

```{r echo=FALSE}
incubation_hatch_raw |>
  group_by(spawning_date) |> 
  summarize(mean_hatch = mean(hatch_percent)) |> 
  ggplot(aes(spawning_date, mean_hatch)) +
  geom_col() +
  theme_minimal()
```

Egg Volume vs. Fry Production (Scatter Plot)

```{r echo=FALSE}
incubation_hatch_raw |>
  ggplot(aes(egg_volume_m_l, total_fry)) +
  geom_point() +
  theme_minimal()
```

Total Eggs ovs Total Fry hatched 
 
```{r echo=FALSE}
incubation_hatch_raw |>
  ggplot(aes(total_eggs, total_fry)) +
  geom_point() +
  theme_minimal()
```

## Historical Larval Sucker Collection Summary (2016–2024)

Historical timeline of the wild larval suckers collected from the lower Williamson River near Chiloquin, Oregon, used for production purposes at the Klamath Falls National Fish Hatchery (KFNFH)

```{r include=FALSE}
table_4_raw <- tables_raw[[5]] |> 
  glimpse()
```

```{r echo=FALSE}
sucker_larval_catch <- table_4_raw |> 
  clean_names() |> 
  slice(-10) |> 
  mutate(year_spring = as.numeric(year_spring),
         number_collected = readr::parse_number(number_collected)) |> 
  glimpse()
```

Annual Larval Collection Trend 

```{r echo=FALSE}
# Collected plotted against Year (Spring)
sucker_larval_catch |> 
  ggplot(aes(x = year_spring, y = number_collected, color = primary_collection_method)) +
  geom_point() +
  theme_minimal()

```

## Larval Early Rearing Mortality and Survival by Tank (FY24)

 - culture_unit: specific tank the fish were placed in (e.g., C3, C4). rows marked "Restock" indicate instances where a culture unit was used for a second or subsequent batch of collected fish
 - collected: initial number of fish placed in the unit (Total: 30,150)
 - stocked: final number of fish transferred or stocked out of the tank 
 - unobserved_mortality_percent: calculated based on the difference between collected fish and the sum of observed mortality and stocked fish 

```{r include=FALSE}
table_5_raw <- tables_raw[[6]] |> 
  clean_names() |> 
  glimpse()
```

```{r echo=FALSE}
rearing_performance <- table_5_raw |> 
  mutate(date= as.Date(date, format = "%m/%d/%Y")) |> 
  slice(-26) |>  #removing total for now
  glimpse()
```

```{r eval=FALSE, include=FALSE}
# Mortality and Loss Breakdown by Culture Unit (Stacked Bar Chart):
# The collected fish count, stacked or broken down by three categories: Stocked, Observed Mortality, and Unobserved Mortality for each Culture Unit.

rearing_performance |> 
  group_by(culture_unit) |> 
  summarise(total_collected = sum(collected)) |> view()
  ggplot()

```

Mortality and Loss Breakdown by Culture Unit

```{r echo=FALSE}
ggplot(rearing_performance, aes(x = culture_unit)) +
  geom_bar(aes(y = stocked, fill = "Stocked"), stat = "identity") +
  geom_bar(aes(y = observed_mortality_percent, fill = "Observed Mortality"), stat = "identity") +
  geom_bar(aes(y = unobserved_mortality_percent, fill = "Unobserved Mortality"), stat = "identity") +
  scale_fill_manual(name = "Category",
                    values = c("Stocked" = "gray",
                               "Observed Mortality" = "black",
                               "Unobserved Mortality" = "salmon")) +
  labs(title = "",
       x = "Culture Unit",
       y = "Value (Count or %)") +
  theme_minimal(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Observed Mortality, Unobserved Mortality, and Survival Over Date

```{r echo=FALSE}
rearing_performance |> 
  ggplot(aes(x = date)) +
  geom_line(aes(y = observed_mortality_percent, color = "Observed Mortality")) +
  geom_line(aes(y = survival_percent, color = "Survival")) +
  geom_line(aes(y = unobserved_mortality_percent, color = "Unobserved Mortality")) +
  scale_color_manual(values = c("Observed Mortality" = "#5DA899",
                                "Unobserved Mortality" = "#2E2585",
                                "Survival" = "#DCCD7D")) +
  theme_minimal()
```

## Pond Monitoring - Spring 2023 – Fall 2023 - Main Growout Season (pre-FY24)

Pond, collection year, stocking date, number of suckers on stocking date, average weight per fish, total weight of all fish in each pond (used to derive feeding amounts), average total length on stocking date (mm, measured or derived from length weight relationships, date harvested, number of suckers on harvest date, growth, and survival for ponds during the production cycle starting in the Spring of 2023 through the Fall of 2023 at the Klamath Falls National Fish Hatchery.

- P-series vs A-series ponds: two sets of outdoor rearing spaces distinguished primarly by size and operational status
- P-Series pond: Small ponds, typically 0.03-acre each, 
  - Primary growout space. 
  - Twenty-two P-series ponds remained operational throughout most of FY24
  
- A-series ponds: Larger ponds, typically 0.25-acre each
  - Demolished early in FY24
  
- "start_weight_fish" (Average Weight Per Fish in Grams): This column records the average weight of a single sucker within a given pond lot at the time the fish were initially stocked into that pond
- "start_wt_g" (Total Weight of All Fish in Grams): This column records the cumulative total weight (biomass) of all the suckers in that specific pond lot at the time they were stocked ("Start Date"). This total weight is the critical metric used by staff to derive daily feeding amounts for the fish in that pond

```{r include=FALSE}
table_6_raw <- tables_raw[[8]] |> 
  row_to_names(row_number = 1) |> 
  clean_names() |> 
  slice(1:25)
```


```{r echo=FALSE}
pond_monitoring <- table_6_raw |> 
  mutate(stocked_date = as.Date(stocked_date, format = "%m/%d/%Y"),
         harvest_date = as.Date(harvest_date, format = "%m/%d/%Y"),
         start_number = as.numeric(start_number),
         start_g_fish = as.numeric(g_fish),
         start_wt_g = as.numeric(g),
         start_wt_lb = as.numeric(lb),
         start_tl_mm = as.numeric(mm),
         end_number = as.numeric(end_number),
         end_g_fish = as.numeric(g_fish_2),
         end_wt_g = as.numeric(g_2),
         end_gt_lb = as.numeric(lb_2),
         end_tl_mm = as.numeric(mm_2),
         days = as.numeric(days),
         months = as.numeric(months),
         growth_mm_day = as.numeric(mm_day),
         weight_gain_lb = as.numeric(lb_3),
         harvest_mortality_number = as.numeric(number),
         survival_percent = as.numeric(percent)) |> 
  glimpse()
```

Growth rate vs start total length in mm

```{r echo=FALSE}
# see growth rates- younger vs older fish?
pond_monitoring |> 
  ggplot(aes(x = start_tl_mm, y = growth_mm_day)) +
  geom_point() +
  theme_minimal()
```

Start and end weight per pond

```{r echo=FALSE}
# see growth rates- younger vs older fish?
# represent the success of the pond as a production unit during that cycle, as start_wt is the total weight used to derive feeding amounts, and the difference between start and end weight shows the total weight gain. The plot could highlight ponds, such as the large A-series ponds (A1, A2, A3), which, despite having been demolished early in FY24, are still represented in this data and likely produced high biomass gains due to their large size
ggplot(pond_monitoring, aes(x = pond)) +
  geom_bar(aes(y = start_wt_g, fill = "start weight (g)"), 
           stat = "identity", position = position_nudge(x = -0.15), width = 0.3) +
  geom_bar(aes(y = end_wt_g, fill = "end weight (g)"), 
           stat = "identity", position = position_nudge(x = 0.15), width = 0.3) +
  scale_fill_manual(values = c("start weight (g)" = "#DCCD7D", "end weight (g)" = "#5DA899")) +
  labs(title = "Start vs End Weight by Pond",
       x = "Pond",
       y = "Weight (g)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

Days in the pond vs growth per day (mm)

```{r echo=FALSE}
# see growth rates- younger vs older fish?
pond_monitoring |> 
  ggplot(aes(x = days, y = growth_mm_day)) +
  geom_point() +
  theme_minimal()
```

## Pond Monitoring - Fall 2023 – Spring 2023 - Main Growout Season (pre-FY24)

This table covers the production cycle starting in the Fall of 2023 through the Spring of 2023 

- question: would this be fall of 2023 only?
- there are a lot of missing values on this table

```{r include=FALSE}
table_7_raw <- tables_raw[[9]] |> 
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names)
    })()
```


```{r echo=FALSE}
pond_monitoring_fall_to_spring <- table_7_raw |> 
  slice(1:16) |> 
  mutate(pond = x1_pond,
         lot = x2_lot,
         stocked_date = as.Date(x3_stocked_date, format = "%m/%d/%Y"),
         harvest_date = as.Date(x9_harvest_date, format = "%m/%d/%Y"),
         start_number = as.numeric(x4_start_number),
         start_g_fish = as.numeric(start_g_fish),
         start_wt_g = as.numeric(start_wt_g),
         start_wt_lb = as.numeric(start_wt_2_lb),
         start_tl_mm = as.numeric(start_tl_mm),
         end_number = as.numeric(x10_end_number),
         end_g_fish = as.numeric(end_g_fish),
         end_wt_g = as.numeric(end_wt_12_g),
         end_gt_lb = as.numeric(end_wt_13_lb),
         end_tl_mm = as.numeric(end_tl_mm),
         days = as.numeric(x15_days),
         months = as.numeric(x16_months),
         growth_mm_day = as.numeric(growth_mm_day),
         weight_gain_lb = as.numeric(weight_gain_lb),
         harvest_mortality_number = as.numeric(harvest_mortality_number),
         survival_percent = as.numeric(survival_percent)) |> 
  select(pond, lot, stocked_date, start_number, start_g_fish, start_wt_g, start_wt_lb, start_tl_mm,
         harvest_date, end_number, end_g_fish, end_wt_g, end_gt_lb, end_tl_mm, days, months, growth_mm_day,
         weight_gain_lb, harvest_mortality_number, survival_percent) |> 
  glimpse()
```

Growth rate vs start total length in mm

```{r echo=FALSE}
# see growth rates- younger vs older fish?
pond_monitoring_fall_to_spring |> 
  ggplot(aes(x = start_tl_mm, y = growth_mm_day)) +
  geom_point() +
  theme_minimal()
```

Start and end weight per stocked date/pond 

```{r echo=FALSE}
ggplot(pond_monitoring_fall_to_spring, aes(x = pond)) +
  geom_bar(aes(y = start_wt_g, fill = "start weight (g)"), 
           stat = "identity", position = position_nudge(x = -0.15), width = 0.3) +
  geom_bar(aes(y = end_wt_g, fill = "end weight (g)"), 
           stat = "identity", position = position_nudge(x = 0.15), width = 0.3) +
  scale_fill_manual(values = c("start weight (g)" = "#DCCD7D", "end weight (g)" = "#5DA899")) +
  labs(title = "",
       x = "Pond",
       y = "Weight (g)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45))
```

```{r echo=FALSE}
pond_monitoring_fall_to_spring |> 
  mutate(stocked_date = as.character(stocked_date)) |>
  ggplot(aes(x = stocked_date)) +
  geom_bar(aes(y = start_wt_g, fill = "start weight (g)"), 
           stat = "identity", position = position_nudge(x = -0.15), width = 0.3) +
  geom_bar(aes(y = end_wt_g, fill = "end weight (g)"), 
           stat = "identity", position = position_nudge(x = 0.15), width = 0.3) +
  scale_fill_manual(values = c("start weight (g)" = "#DCCD7D", "end weight (g)" = "#5DA899")) +
  labs(title = "",
       x = "Stocked Date",
       y = "Weight (g)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Days in the pond vs growth per day (mm)

```{r echo=FALSE}
# see growth rates- younger vs older fish?
pond_monitoring_fall_to_spring |> 
  ggplot(aes(x = days, y = growth_mm_day)) +
  geom_point() +
  theme_minimal()
```

## Pond Monitoring - Fall 2023 or Spring 2024 through Fall 2024

Subsequent growing season 2024 (Fall 2023/Spring 2024 through Fall 2024.
Production cycle starting in the Fall of 2023 or Spring of 2024 through the Fall of 2024 at the Klamath Falls National Fish Hatchery. Please note that Fall inventory work was not completed for most ponds prior to the close of FY 2024.


```{r include=FALSE}
table_8_raw <- tables_raw[[10]] |> 
  clean_names() |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names)
    })()
```


```{r echo=FALSE}
glimpse(table_8_raw)
pond_monitoring_fall_to_fall <- table_8_raw |> 
  # slice(1:16) |> 
  mutate(pond = x1_pond,
         lot = x2_lot,
         stocked_date = as.Date(x3_stocked_date, format = "%m/%d/%Y"),
         harvest_date = as.Date(x9_harvest_date, format = "%m/%d/%Y"),
         start_number = as.numeric(x4_start_number),
         start_g_fish = as.numeric(start_g_fish),
         start_wt_g = as.numeric(start_wt_g),
         start_wt_lb = as.numeric(start_wt_2_lb),
         start_tl_mm = as.numeric(start_tl_mm),
         end_number = as.numeric(x10_end_number),
         end_g_fish = as.numeric(end_g_fish),
         end_wt_g = as.numeric(end_wt_12_g),
         end_gt_lb = as.numeric(end_wt_13_lb),
         end_tl_mm = as.numeric(end_tl_mm),
         days = as.numeric(x15_days),
         months = as.numeric(x16_months),
         growth_mm_day = as.numeric(growth_mm_day),
         weight_gain_lb = as.numeric(weight_gain_lb),
         harvest_mortality_number = as.numeric(harvest_mortality_number),
         survival_percent = as.numeric(survival_percent)) |> 
  select(pond, lot, stocked_date, start_number, start_g_fish, start_wt_g, start_wt_lb, start_tl_mm,
         harvest_date, end_number, end_g_fish, end_wt_g, end_gt_lb, end_tl_mm, days, months, growth_mm_day,
         weight_gain_lb, harvest_mortality_number, survival_percent) |> 
  glimpse()
```

## Sucker Pen survival 

- Each operation year is divided by the different fish pens (UKLN, UKLS, UKLW, UKLE) and the year the fish were captured as larval fish from the Williamson River (CY). 
- CY: captured year
- The average total length (TL) in millimeters (mm): to show survival of each operation year.
- Operation year 2020 is an outlier year as fish were not stocked until fall and only held for two months
- no production fish were stocked in 2024 as the future of using the UKL net pen to grow suckers is being assessed by USFWS.


```{r include=FALSE}
table_9_raw <- tables_raw[[11]] |> 
  clean_names() |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-2, ] |>
      setNames(new_names)
    })()
```


```{r echo=FALSE}
pen_survival <- table_9_raw |> 
  slice(2:12) |> 
  separate(stock_tl_number, into = c("stock_tl_mm", "number_stocked"), sep = "\\s+", remove = TRUE) |> 
  mutate(operation_year = as.numeric(operation),
         captured_year = as.numeric(x2_fish_cy),
         pen_id = x3_pen_id,
         stock_tl_mm = as.numeric(stock_tl_mm),
         number_stocked = as.numeric(number_stocked),
         harvest_tl_mm = as.numeric(harvest_tl),
         number_harvested = as.numeric(number),
         survival_percentage = as.numeric(x8_survival_percent)) |> 
  select(operation_year, captured_year, pen_id, stock_tl_mm, number_stocked, harvest_tl_mm,
         number_harvested, survival_percentage) |> 
  glimpse()

```

Survival Percentage by Pen and Year

```{r echo=FALSE}
ggplot(pen_survival, aes(x = factor(operation_year), y = survival_percentage, fill = pen_id)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  # labs(x = "Operation Year",
  #   y = "Survival Percentage",
  #   fill = "Pen ID",
  #   title = "Survival Percentage by Pen and Year"
  # ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

