---
title: "USFWS Klamath Falls National Fish Hatchery Reports"
author: "Badhia Yunes Katz (FlowWest)"
date: "`r Sys.Date()`"
output: html_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>", 
  fig.width=8, fig.height=5)
library(tidyverse)
library(lubridate)
library(tabulapdf)
library(janitor)
library(dplyr)
library(tidyr)
library(readr)
```

The tables in the Klamath Falls National Fish Hatchery (KFNFH) annual reports are designed to provide detailed documentation of operations and results. These reports were prepared by Prepared by the Klamath Falls National Fish Hatchery.

This markdown will extract and explore the tables on the hatchery table reports. 

It will identify any questions/clarification we might need as we go through the reports.

**Link to Sources**

## Annual Report for Fiscal Year: {.tabset}

### 2021 {.tabset}


#### Table 1

Collection and release information for the Sucker Assisted Rearing Program (SARP) since its inception in 2016 by fiscal years. This table tracks key metrics over time, including Larvae Collected, SARP Release numbers, and Salvage Release numbers.

```{r echo=FALSE}
# TODO tables with historical data - for example table 1. Just pull the most up to date one. In this case 2024
# extract tables
pdf_path_2021 <- "data-raw/fisheries/hatcheries/20240912_FY2021_KFNFH_Annual_Report_Final.pdf"
tables_2021 <- extract_tables(pdf_path_2021, method = "stream", output = "tibble")

table_1 <- tables_2021[[1]] |>
  clean_names()

# fixing so col name grabs the first row too
table_2021_clean <- tables_2021[[1]] |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names) |>
      mutate(across(-1, ~parse_number(.x)))
    })()

knitr::kable(table_2021_clean)

```

#### Table 2

**Summary of Lost River Sucker (LRS), Shortnose Sucker (SNS), and Klamath Largescale Sucker (KLS) stocked during the Fiscal Year 2020 (October 1, 2020, to September 30, 2021).** This table breaks down fish stocking activities by date, species, Lot #, weight, density (#/lb.), Total Length (TL) in inches and mm, the ponds the fish originated from, and the Stocking Location.

*(showing the first 10 first rows only)*
```{r echo=FALSE}
table_2 <- tables_2021[[2]] |>
  clean_names() |> 
  mutate(is_real_date = grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", date), # dates (mm/dd/yyyy)
         date_lower = tolower(date),  # full month names
         is_month = date_lower %in% tolower(month.name),
         is_grand_total = grepl("^grand\\s*totals?$", date_lower, ignore.case = TRUE), #"Grand Totals" 
         date_anchor = case_when(is_real_date ~ date,
                                 is_month ~ date,
                                 is_grand_total ~ date,
                                 TRUE ~ NA_character_)) |> 
  fill(date_anchor, .direction = "down") |> 
  mutate(date = if_else(is.na(date), date_anchor, date)) |> 
  select(-is_real_date, -is_month, -is_grand_total, -date_lower, -date_anchor) 

# split to keep the totals on a separate table
table_2_dates <- table_2 |>
  filter(grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", date))

table_2_summary <- table_2 |>
  filter(!grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", date)) |>
  select(-c(from_ponds, stocking_location)) 

knitr::kable(head(table_2_dates, 10))
knitr::kable(head(table_2_summary, 10))  

```

#### Table 3

*Fall 2020 production summary.* This table summarizes activities during the fall of 2020, tracking fish cohorts (Lot #) by Harvest Date and Pond, noting inventory numbers (Paper # and Actual #), survival rates, the number and percentage of fish stocked, the number and percentage of fish graded as "Short," the number of fish designated as broodstock (# Brood), and the disposition of the fish

```{r echo=FALSE}
table_3 <- tables_2021[[3]] |>
  clean_names() 
  
table_3_clean <- table_3 |> 
  separate(harvest_date_pond, into = c("harvest_date", "pond"), sep = "\\s+", remove = TRUE) |> 
    mutate(harvest_lower = tolower(harvest_date),
           is_month_or_season = harvest_lower %in% c(tolower(month.name), "fall"),
           anchor = if_else(is_month_or_season, harvest_date, NA_character_)) |>
  fill(anchor, .direction = "down") |>
  mutate(harvest_date = if_else(is.na(harvest_date), anchor, harvest_date)) |>
  select(-harvest_lower, -is_month_or_season, -anchor, -x3) |> 
  separate(number_stocked_percent_stocked,
           into = c("number_stocked", "percent_stocked"),
           sep = "\\s+",
           fill = "right",
           extra = "drop",
           convert = TRUE) 

# split into two tables - separating totals to another table
fall_production_summary <- table_3_clean |> 
  filter(grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", harvest_date)) |> 
  mutate(harvest_date = as.Date(harvest_date, format = "%m/%d/%Y"))

fall_production_summary_totals <- table_3_clean |> 
  filter(!grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", harvest_date)) |> 
  mutate(harvest_date = ifelse(!is.na(pond), paste(harvest_date, pond), harvest_date)) |> 
  select(-c(pond, disposition, tl_mm_7, tl_mm_9, tl_mm_12)) |> 
  mutate(harvest_date = case_when(harvest_date == "September" ~ "September Totals",
                                  harvest_date == "November" ~ "November Totals",
                                  harvest_date == "December" ~ "December Totals",
                                  TRUE ~ harvest_date))

knitr::kable(head(fall_production_summary, 10))
knitr::kable(head(fall_production_summary_totals, 10))  
```

#### Table 4

Spring 2021 production summary. 

```{r echo=FALSE}
table_4 <- tables_2021[[4]] |>
  clean_names() |> 
    separate(harvest_date_pond, into = c("harvest_date", "pond"), sep = "\\s+", remove = TRUE) |> 
    mutate(harvest_lower = tolower(harvest_date),
           is_month_or_season = harvest_lower %in% c(tolower(month.name), "spring"),
           anchor = if_else(is_month_or_season, harvest_date, NA_character_)) |>
  fill(anchor, .direction = "down") |>
  mutate(harvest_date = if_else(is.na(harvest_date), anchor, harvest_date)) |>
  select(-harvest_lower, -is_month_or_season, -anchor, -x2, -tl_mm_10, -number_short_percent_short, -x12, -tl_mm_13, -number_brood) |>
  separate(number_stocked_percent_stocked,
           into = c("number_stocked", "percent_stocked"),
           sep = "\\s+",
           fill = "right",
           extra = "drop",
           convert = TRUE) 

# split into two tables - separating totals to another table
spring_production_summary <- table_4 |> 
  filter(grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", harvest_date)) |> 
  mutate(harvest_date = as.Date(harvest_date, format = "%m/%d/%Y"))

spring_production_summary_totals <- table_4 |> 
  filter(!grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", harvest_date)) |> 
  mutate(harvest_date = ifelse(!is.na(pond), paste(harvest_date, pond), harvest_date)) |>
  select(-disposition_stocked_to, -tl_mm_8, -pond) |>
 mutate(harvest_date = case_when(
   str_detect(harvest_date, regex("February", ignore_case = TRUE)) ~ "February 2021 Totals",
   str_detect(harvest_date, regex("March", ignore_case = TRUE)) ~ "March 2021 Totals",
   str_detect(harvest_date, regex("April", ignore_case = TRUE)) ~ "April 2021 Totals",
   str_detect(harvest_date, regex("May", ignore_case = TRUE)) ~ "May 2021 Totals",
   str_detect(harvest_date, regex("Spring", ignore_case = TRUE)) ~ "Spring 2021 Overall",
   TRUE ~ harvest_date))
  
knitr::kable(head(fall_production_summary, 10))
knitr::kable(head(fall_production_summary_totals, 10))  
```

#### Table 5 

Summary of PIT Tag Study Pond stockings in Replicate Block 1 during week of March 22, 2021.

```{r echo=FALSE}
table_5 <- tables_2021[[5]] |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    
    x[-1, ] |>
      setNames(new_names) |>
      mutate(
        across(
          # apply parse_number to all except first column and tagged_vs_untagged
          -c(1, tagged_vs_untagged),
          ~parse_number(.x)
        )
      )
  })()

knitr::kable(table_5)
```

#### Table 6 

Summary of PIT Tag Study Pond stockings in Replicate Block 2 during week of March 29, 2021. 

```{r echo=FALSE}
table_6 <- tables_2021[[6]] |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    
    x[-1, ] |>
      setNames(new_names) |>
      mutate(
        across(
          # apply parse_number to all except first column and tagged_vs_untagged
          -c(1, tagged_vs_untagged),
          ~parse_number(.x)
        )
      )
  })()

knitr::kable(table_6)
```

#### Table 7

Summary of PIT Tag Study Pond stockings in Replicate Block 3 during week of April 5, 2021.

```{r echo=FALSE}
table_7 <- tables_2021[[7]] |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    
    x[-1, ] |>
      setNames(new_names) |>
      mutate(
        across(
          # apply parse_number to all except first column and tagged_vs_untagged
          -c(1, tagged_vs_untagged),
          ~parse_number(.x)
        )
      )
  })()

knitr::kable(table_7)
```

### 2022 {.tabset}

Some of the tables are not being read for now #TODO figure out approach

#### Table 1

Collection and release information for the Klamath Falls National Fish Hatchery since its 
inception in 2016 by fiscal years; federal fiscal years run from October 1st through the following 
September 30th of each year. During fiscal year 2022, 12,768 production fish and 4,655 fingerlings were released, and 19,993 experimentally hatched fry were transferred to another hatchery, totaling 37,416 produced in FY22. 

```{r echo=FALSE}
# extract tables
pdf_path_2022 <- "data-raw/fisheries/hatcheries/20240830_FY2022 KFNFH Annual Report_Final.pdf"
tables_2022 <- extract_tables(pdf_path_2022, method = "stream", output = "tibble")

table_1_2 <- tables_2022[[2]] |>
  clean_names()

table_1_2_clean <- table_1_2 |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names)
    })()

collection_release <- table_1_2_clean |> 
  rename(tl_1_mm = tl_4_mm,
         tl_2_mm = tl_6_mm,
         tl_3_mm = tl_8_mm) 

knitr::kable(collection_release)
```

#### Table 2

Inventory, growth, and survival for ponds during the production cycle starting in the 
Spring of 2021 through the Fall of 2021 at the Klamath Falls National Fish Hatchery.

```{r echo=FALSE}
table_2_2 <- tables_2022[[3]] |>
  clean_names()

table_2_2_clean <- table_2_2 |>
  separate(growth_mm_day_total_growth_mm,
           into = c("growth_mm_day", "total_growth_mm"),
           sep = "\\s+",
           fill = "right",
           extra = "drop",
           convert = TRUE)
  
inventory_growth_spring_fall <- table_2_2_clean |> 
  slice(1:46) |> 
  mutate(stocked_to_pond  = as.Date(stocked_to_pond, format = "%m/%d/%Y"),
         harvest_date = as.Date(harvest_date, format = "%m/%d/%Y")) 

inventory_growth_spring_fall_summary <- table_2_2_clean |> 
  slice(47:52) |>
  select(-c(pond, stocked_to_pond, start_tl_mm, harvest_date, end_tl_mm, total_days, total_growth_mm, growth_mm_day))

knitr::kable(inventory_growth_spring_fall)
knitr::kable(inventory_growth_spring_fall_summary)
```

#### Table 5

Number of suckers raised and released by the Klamath Falls National Fish Hatchery during 
fiscal year 2022, spanning from October 1, 2021 to September 30, 2022. 

```{r echo=FALSE}
table_3_2 <- tables_2022[[4]] |>
  clean_names()

table_3_2_clean <- table_3_2 |> 
   mutate(is_real_date = grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", date),
          date_lower = tolower(date),
          is_month = date_lower %in% tolower(month.name),
          is_period = str_detect(date_lower,
                                 "^(fall|spring|summer|winter)(\\s+of)?\\s+\\d{4}$") |
            str_detect(date_lower, "fiscal year\\s*\\d{4}$") |
            str_detect(date_lower, "all of fiscal year\\s*\\d{4}$"),
          is_grand_total = grepl("^grand\\s*totals?$", date_lower, ignore.case = TRUE),
          date_anchor = case_when(is_real_date ~ date,
                                  is_month ~ date,
                                  is_period ~ date,
                                  is_grand_total ~ date,
                                  TRUE ~ NA_character_)) |> 
  fill(date_anchor, .direction = "down") |> 
  mutate(date = if_else(is.na(date), date_anchor, date)) |>
  select(-is_real_date, -is_month, -is_period, -is_grand_total, -date_lower, -date_anchor) |>
  glimpse()

# split to keep the totals on a separate table
suckers_raised <- table_3_2_clean |>
  filter(grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", date)) |> 
  mutate(date = as.Date(date, format = "%m/%d/%Y")) |> glimpse()

suckers_raised_summary <- table_3_2_clean |>
  filter(!grepl("^\\d{1,2}/\\d{1,2}/\\d{4}$", date)) |>
  select(-c(stocking_location, actual_tl_mm)) |> glimpse()


knitr::kable(suckers_raised)
knitr::kable(suckers_raised_summary)
```

#### Table 6

Number of individual passive integrated transponder (PIT) tagged suckers 
released by the Klamath Falls National Fish Hatchery and detected on remote arrays within the Upper 
Klamath Lake Basin for each release and detection year through April of 2022 as the last data release. 
```{r echo=FALSE}
table_4_2 <- tables_2022[[5]] |>
  clean_names()

table_4_2_clean <- table_4_2 |> 
separate(detection_year,
           into = c("2014", "2015"),
           sep = "\\s+",
           fill = "right",
           extra = "drop",
           convert = TRUE) %>%
  {
    header <- as.character(.[1, ])
    fallback <- as.character(.[2, ])
    names(.) <- ifelse(is.na(header) | header == "", fallback, header)
    .[-1, ]   # drop the header row
  } |> 
  slice(-1) |> 
  # pivot this table to assign release year and detection year
  pivot_longer(cols = matches("^\\d{4}$"),
               names_to = "Detection_Year",
               values_to = "Count") |> 
  mutate(release_year = as.integer(`Release Year`),
         detection_year = as.integer(Detection_Year),
         count = as.numeric(Count)) |> 
  clean_names()

pit_tag <- table_4_2_clean |> 
  select(release_year, detection_year, count) 


pit_tag_totals <- table_4_2_clean |> 
  select(release_year, release_year_total) 

knitr::kable(pit_tag)
knitr::kable(pit_tag_totals)
```


```{r eval=FALSE, include=FALSE}
# these two approaches are failing to read the tables
#TODO find a different

# usging tabula
# Extract only pages 10 and 11 - for some reason these two tables are not being read
# tables_2022_p10_11 <- extract_tables(file = pdf_path_2022,
#                                      method = "stream",
#                                      pages = 10:11,
#                                      output = "tibble")

# using pdftools
# library(pdftools)
# page_text <- pdf_text(pdf_path_2022)[10:11]
# 
# cat(page_text[1])  # page 10
# cat(page_text[2])  # page 11

```

### 2023 {.tabset}


#### Table 1

Collections and releases from the Klamath Falls National Fish Hatchery since its inception in 
2016, by federal fiscal years, running from October 1 through the following September 30. All total 
lengths (TL) and standard lengths (SL) reported are averages based on length-weight conversions from 
cumulative total weights of each type, production, fingerling, fry, and salvage fish for each year. The line for fiscal year 2024, which includes fish stocked between October 1 through December 31, 2023, is currently in progress at the time of writing and is included so the most current information is available on releases.  

```{r echo=FALSE}
# extract tables
pdf_path_2023 <- "data-raw/fisheries/hatcheries/20240801_FY2023 KFNFH Annual Report_Final.pdf"
tables_2023 <- extract_tables(pdf_path_2023, method = "stream", output = "tibble")

table_1_3 <- tables_2023[[2]] |>
  clean_names()

table_1_3_clean <- table_1_3 |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names)
    })()

collection_release_2023 <- table_1_3_clean |> 
  rename(tl_1_mm = tl_4_mm,
         tl_2_mm = tl_6_mm,
         tl_3_mm = tl_8_mm) 

knitr::kable(collection_release_2023)
```


#### Table 2

Summary of the wild East Side Springs Lost River Suckers handled during spawning activities 
from Upper Klamath Lake in the spring of 2023. AFTC stands for Abernathy Fish Technology Center.

```{r echo=FALSE}
table_2_3 <- tables_2023[[3]] |>
  clean_names()

summary_esp_lrs <- table_2_3 |> 
  separate(
    col = pit_tag_last_5_sex_fork_length_atfc_genetic_id,
    into = c("PIT_TAG_LAST_5", "SEX", "FORK_LENGTH", "ATFC_GENETIC_ID"),
    sep = "\\s+",     # split on one or more spaces
    remove = TRUE) |> 
  separate(
    col = date_spring_name,
    into = c("date", "spring_name"),
    sep = "\\s+",     # split on one or more spaces
    remove = TRUE) |> 
  select(-c(x2, x4, x5)) |> 
  clean_names() |> 
  mutate(atfc_genetic_id = case_when(atfc_genetic_id == "LABELED" ~ "3960-019",
                                     atfc_genetic_id == "(DID" ~ " 3960-028",
                                     TRUE ~ atfc_genetic_id),
         date = as.Date(date, format = "%m/%d/%Y"),
         pit_tag_last_5 = case_when(pit_tag_last_5 %in% c("SAMPLE", "TWO") ~ NA_character_,
                                    TRUE ~ pit_tag_last_5),
         sex = case_when(sex %in% c("NO", "FISH") ~ NA_character_,
                                    TRUE ~ sex),
         fork_length = as.numeric(fork_length)) 

knitr::kable(head(summary_esp_lrs, 10)) 
```

#### Table 3

Summary of the wild East Side Springs Lost River Suckers handled during spawning activities 
from Upper Klamath Lake in the spring of 2023. AFTC stands for Abernathy Fish Technology Center.

```{r echo=FALSE}
table_3_4 <- tables_2023[[4]] |>
  clean_names() |> 
  rename(observed = observed_4,
         observed_2 = observed_7) |> glimpse()

table_3_4_clean <- table_3_4 %>% 
  {
    if (nrow(.) >= 2) {
      header <- purrr::map_chr(seq_along(.), function(i) {
        parts <- c(
          names(.)[i],
          as.character(.[1, i]),
          as.character(.[2, i])
        )
        parts <- parts[!is.na(parts) & parts != ""]
        trimws(paste(parts, collapse = " "))
      }) %>%
        gsub("\\s+", " ", .)

      names(.) <- header
      .[-c(1, 2), ] 
    } else {
      .
    }
  }|> 
  clean_names()

table_3_4_clean <- table_3_4_clean |> 
    mutate(collection_date = as.Date(collection_date, format = "%m/%d/%Y"),
           collected_number = readr::parse_number(collected_number),
           collected_number = as.numeric(collected_number),
           observed_mortality_number = readr::parse_number(observed_mortality_number),
           observed_mortality_number = as.numeric(observed_mortality_number),
           stocked_number = readr::parse_number(stocked_number),
           stocked_number = as.numeric(stocked_number),
           overall_survival_percent = as.numeric(overall_survival_percent),
           observed_2_mortality_percent = as.numeric(observed_2_mortality_percent),
           unobserved_mortality_percent = as.numeric(unobserved_mortality_percent))

initial_stock <- table_3_4_clean |> 
  slice(1:14) |> 
  mutate(collection_date = as.Date(collection_date, format = "%m/%d/%Y"),
         collected_number = as.numeric(collected_number),
         observed_mortality_number = as.numeric(observed_mortality_number),
         stocked_number = as.numeric(stocked_number),
         overall_survival_percent = as.numeric(overall_survival_percent),
         observed_2_mortality_percent = as.numeric(observed_2_mortality_percent),
         unobserved_mortality_percent = as.numeric(unobserved_mortality_percent)) 

initial_stock_totals <- table_3_4_clean |> 
  slice(15) 

knitr::kable(head(initial_stock, 10)) 
knitr::kable(head(initial_stock_totals, 10)) 
```

#### Table 4

 Historical catches of larval suckers from the lower Williamson River, near Chiloquin, Oregon. (*) 
Note that 2023 was the first year when collections were limited in the number of fry available during the 
collection season. 

```{r echo=FALSE}
table_4_4 <- tables_2023[[5]] |>
  clean_names() 

historical_catches <- table_4_4 |> 
  mutate(number_collected = readr::parse_number(number_collected),
         year_spring = as.numeric(year_spring)) |> 
  slice(1:8) 

historical_catches_total <- table_4_4 |> 
  mutate(number_collected = readr::parse_number(number_collected)) |> 
  select(-primary_collection_method) |> 
  slice(9) 

knitr::kable(head(historical_catches, 10)) 
knitr::kable(head(historical_catches_total, 10)) 
```

#### Table 5

Pond, collection year, stocking date, number of suckers on stocking date, average weight per fish, total weight of all fish in each pond (used to derive feeding amounts), average total length on stocking date (mm, measured or derived from length weight relationships, date harvested, number of suckers on harvest date, growth, and survival for ponds during the production cycle starting in the Spring of 2022 through the Fall of 2022 at the Klamath Falls National Fish Hatchery. The last seven rows in the table are summary numbers for each year class on station and the overall for the harvest season. 

This table details the pond performance during the production cycle starting in the Spring of 2022 through the Fall of 2022

```{r echo=FALSE}
# similar than table 2 from 2022
table_5_4 <- tables_2023[[6]] |>
  clean_names() |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names)
    })()

table_5_4_clean <- table_5_4 |> 
  mutate(year_split = str_split(lot, "/")) |> #creating two rows for lot 2021/2019
  unnest(year_split) |> 
  mutate(lot = as.integer(year_split)) |> 
  select(-year_split) 

inventory_clean_2023 <- table_5_4_clean |> 
  mutate(stocked = as.Date(stocked, format = "%m/%d/%Y"),
         total_number_1 = as.numeric(total_4_number),
         average_g_fish = as.numeric(average_5_g_fish),
         weight_g_1 = as.numeric(weight_6_g),
         weight_lbs_1 = as.numeric(weight_7_lbs),
         tl_mm_1 = as.numeric(tl_8_mm),
         harvested = as.Date(harvested, format = "%m/%d/%Y"),
         total_number_2 = as.numeric(total_10_number),
         average_g_fish_2 = as.numeric(average_11_g_fish),
         weight_g_2 = as.numeric(weight_12_g),
         weight_lbs_2 = as.numeric(weight_13_lbs),
         tl_mm_2 = as.numeric(tl_14_mm),
         growth_mm_day = as.numeric(growth_17_mm_day),
         growth_g_fish_mon = as.numeric(growth_18_g_fish_mon),
         weight_gain_lbs = as.numeric(weight_gain_lbs),
         harvest_morts_number = as.numeric(harvest_morts_number),
         survival_percent = as.numeric(survival_percent)) |> 
  select(pond, lot, stocked, total_number_1, average_g_fish, weight_g_1, weight_lbs_1, tl_mm_1,
         harvested, total_number_2, average_g_fish_2, weight_g_2, weight_lbs_2, tl_mm_2, days, months, growth_mm_day,
         growth_g_fish_mon, weight_gain_lbs, harvest_morts_number, survival_percent) 

inventory_2023 <- inventory_clean_2023 |> 
  slice(1:41) 

inventory_2023_summary <- inventory_clean_2023 |> 
  slice(42:48) |> 
  mutate(lot = case_when(is.na(lot) ~ "overall",
                         TRUE ~ as.character(lot))) |> 
  select(lot, total_number_1, total_number_2, harvest_morts_number, survival_percent)

knitr::kable(head(inventory_2023, 10)) 
knitr::kable(head(inventory_2023_summary, 10)) 
```

#### Table 6

Pond, collection year, stocking date, number of suckers on stocking date, average weight per fish, total weight of all fish in each pond (used to derive feeding amounts), average total length on stocking date (mm, measured or derived from length weight relationships, date harvested, number of suckers on harvest date, growth, and survival for ponds during the production cycle starting in the Summer/Fall of 2022 through the Spring of 2023 at the Klamath Falls National Fish Hatchery. The last eight rows in the table are summary numbers for each year class on station and the overall for the harvest season.  
This table summarizes the pond performance during the production cycle starting in the Summer/Fall of 2022 through the Spring of 2023

```{r echo=FALSE}
# this table can probably be mixed with table 5
table_6_4 <- tables_2023[[7]] |>
  clean_names() |>
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names)
    })()

table_6_4_clean <- table_6_4 |> 
  mutate(year_split = str_split(lot, "-")) |> #creating two rows for lot 2021/2019
  unnest(year_split) |> 
  mutate(lot = as.integer(year_split)) |> 
  select(-year_split) |>
  rename(tl_8_mm = start_tl_8_mm, # renaming just to match to last table & be able to use same code
         tl_14_mm = start_tl_14_mm) 


inventory_clean_2022_2023 <- table_6_4_clean |> 
  mutate(stocked = as.Date(stocked, format = "%m/%d/%Y"),
         total_number_1 = as.numeric(total_4_number),
         average_g_fish = as.numeric(average_5_g_fish),
         weight_g_1 = as.numeric(weight_6_g),
         weight_lbs_1 = as.numeric(weight_7_lbs),
         tl_mm_1 = as.numeric(tl_8_mm),
         harvested = as.Date(harvested, format = "%m/%d/%Y"),
         total_number_2 = as.numeric(total_10_number),
         average_g_fish_2 = as.numeric(average_11_g_fish),
         weight_g_2 = as.numeric(weight_12_g),
         weight_lbs_2 = as.numeric(weight_13_lbs),
         tl_mm_2 = as.numeric(tl_14_mm),
         growth_mm_day = as.numeric(growth_17_mm_day),
         growth_g_fish_mon = as.numeric(growth_18_g_fish_mon),
         weight_gain_lbs = as.numeric(weight_gain_lbs),
         harvest_morts_number = as.numeric(harvest_morts_number),
         survival_percent = as.numeric(survival_percent)) |> 
  select(pond, lot, stocked, total_number_1, average_g_fish, weight_g_1, weight_lbs_1, tl_mm_1,
         harvested, total_number_2, average_g_fish_2, weight_g_2, weight_lbs_2, tl_mm_2, days, months, growth_mm_day,
         growth_g_fish_mon, weight_gain_lbs, harvest_morts_number, survival_percent)

inventory_2022_2023 <- inventory_clean_2022_2023 |> 
  slice(1:25)

inventory_2022_2023_summary <- inventory_clean_2022_2023 |> 
  slice(26:33) |>
  mutate(lot = case_when(total_number_1 == 1090 ~ "mixed",
                         total_number_1 == 36058 ~ "overall",
                         TRUE ~ as.character(lot))) |>
  select(lot, total_number_1, total_number_2, harvest_morts_number, survival_percent) 

knitr::kable(head(inventory_2022_2023, 10)) 
knitr::kable(head(inventory_2022_2023_summary, 10)) 
```


#### Table 7

Pond, collection year, stocking date, number of suckers on stocking date, average weight per fish, total weight of all fish in each pond (used to derive feeding amounts), average total length on stocking date (mm, measured or derived from length weight relationships, date harvested, number of suckers on harvest date, growth, and survival for ponds during the production cycle starting in the Spring of 2023 through the Fall of 2023 at the Klamath Falls National Fish Hatchery. The last nine rows in the table are summary numbers for each year class on station and the overall for the harvest season.
This table covers the pond performance during the production cycle starting in the Spring of 2023 through the Fall of 2023. Notably, Table 7 also includes a column for Feed Conversion (Feed Conv.) data, which is not explicitly listed in the column descriptions of the prior two tables.

```{r echo=FALSE}
# this table can probably be mixed with table 5
table_7_4 <- tables_2023[[8]] |>
  clean_names() |> 
  (\(x) {
    new_names <- paste(names(x), x[1, ], sep = "_") |>
      gsub("_NA", "", x = _) |>
      make_clean_names()
    x[-1, ] |>
      setNames(new_names)
    })()

#TODO use similar togic thN Table 6 and 5
```


#### Table 8

Historical survival rates from the Upper Klamath Lake net pens since the initial deployments, first at Rattlesnake Point, in Upper Klamath Lake in 2019-2020 and also at Four Mile Canal, in Agency Lake, near Chiloquin, Oregon in 2021.

```{r echo=FALSE}
# this table can probably be mixed with table 5
table_8_4 <- tables_2023[[9]] |>
  clean_names() 

historical_survival_ukl_clean <- table_8_4 |>
  rename(tl_mm = tl_mm_5,
         tl_mm_2 = tl_mm_7)

historical_survival_ukl <- historical_survival_ukl_clean |> 
  slice(1:11) |> 
  mutate(operation_year = as.numeric(operation_year)) 

historical_survival_ukl_summary <- historical_survival_ukl_clean |> 
  slice(12) |>
  select(operation_year, number_stocked, number_harvested, survival_percent) 

knitr::kable(historical_survival_ukl)
knitr::kable(historical_survival_ukl_summary)
```

#### Table 9

Historical survival rates from the Gerber Reservoir net pens since the initial deployments, near Bonanza, Oregon.

```{r echo=FALSE}
# this table can probably be mixed with table 5
# TODO confirm how we wan tot treat 2022 "Low Water - Not Operational"
table_9_4 <- tables_2023[[10]] |>
  clean_names() |>
  separate(
    col = number_stocked_tl_mm ,
    into = c("number_stocked", "tl_mm"),
    sep = "\\s+",     # split on one or more spaces
    remove = TRUE) |> 
  mutate(number_stocked = as.numeric(number_stocked),
         tl_mm = as.numeric(tl_mm)) |> 
  select(-x6) 

historical_survival_gr <- table_9_4 |>
  slice(1:7) |> 
  mutate(operation_year = as.numeric(operation_year)) 

# TODO this can be combined with historical_survival_ukl_summary
historical_survival_gr_summary <- table_9_4 |>
  slice(8) |>
  select(-year_class, -pen_id, -tl_mm) 
  
knitr::kable(historical_survival_gr)
knitr::kable(historical_survival_gr_summary)
```

#### Table 10

Summary of all fish distributions for repatriation or transfer from the Klamath Falls National Fish Hatchery during the fiscal year of 2023, dating from October 1, 2022 through September 30, 2023.

```{r echo=FALSE}
# TODO this table can probably be mixed with table 11
# we can add column called fiscal_year - this data would be 2023, next table would be 2024
table_10_4 <- tables_2023[[11]] |>
  clean_names() 

fish_dist_2023 <- table_10_4 |> 
  slice(1:63) |> 
  mutate(date = as.Date(date, format = "%m/%d/%Y")) 

#TODO clean last part of table
fish_dist_2023_summary <- table_10_4 |> 
  slice(64:101) |> 
  select(-stocking_location, -from_ponds, -actual_tl_mm) 

knitr::kable(fish_dist_2023)
# knitr::kable(fish_dist_2023_summary)
```


#### Table 11

Summary of all fish distributions for repatriation or transfer from the Klamath Falls National Fish Hatchery during the fiscal year of 2024, dating from October 1, 2023 through September 30, 2024.

```{r echo=FALSE}
# this table can probably be mixed with table 10
table_11_4 <- tables_2023[[12]] |>
  clean_names() 

fish_dist_2024 <- table_11_4 |> 
  slice(1:20) |> 
  mutate(date = as.Date(date, format = "%m/%d/%Y")) |> glimpse()


#TODO figure out how to handle date col
fish_dist_2024_summary <- table_11_4 |> 
  slice(21:30) |> 
  select(-stocking_location, -from_ponds, -actual_tl_mm) 

knitr::kable(fish_dist_2024)
# knitr::kable(fish_dist_2024_summary)
```

#### Table 12

Nutrition of the four commercial diets used in the long term feed diet trial.   

```{r eval=FALSE, include=FALSE}
table_12_4 <- tables_2023[[13]] 

table_12_4_clean <- table_12_4 |>
  rename(Diet = nutrition_comparison) |>
  mutate(Diet = ifelse(grepl("%", Diet), NA, Diet)) |>
  fill(Diet, .direction = "down") |>
  filter(!Diet %in% c("Diet  Protein", "%")) |>
  rename(`Protein %` = Diet,  
         `Fat %` = x2,
         `Fiber %` = x3,
         `Ash %` = x4) |>
  mutate(across(everything(), \(x) trimws(x))) |>
  mutate(`Diet` = case_when(
    grepl("Razorback", `Protein %`) ~ "Razorback",
    grepl("June", `Protein %`) ~ "June",
    grepl("Shri", `Protein %`) ~ "Shrimp",
    TRUE ~ `Protein %`)) |>
  mutate(`Protein %` = readr::parse_number(`Protein %`)) |>
  select(Diet, `Protein %`, `Fat %`, `Fiber %`, `Ash %`)

fish_diet <- table_12_4_clean |> 
  clean_names()
  
knitr::kable(fish_diet)
```

#### Table 13

Comparison of performance between the 4 diets at the completion of the fingerling stage. Included are the mean and standard deviation for total length and weight, and the condition factor and percent survival are also included.  

```{r echo=FALSE}
table_13_4 <- tables_2023[[14]] |> 
  clean_names() |> 
  glimpse()

```

#### Table 14

Comparison of performance between the 4 diets at the completion of the juvenile 
stage. Included are the mean and standard deviation for total length and weight, and the condition factor and percent survival are also included. 

```{r echo=FALSE}
# end results of juvenile stage
table_14_4 <- tables_2023[[15]] |> 
  clean_names() |> 
  row_to_names(row_number = 1) |> 
  glimpse() #TODO fix condition factor
```

#### Table 15

Percent of population at the end of each stage that had observed deformities. Scale loss and hemorrhaging not included in analysis.

```{r echo=FALSE}
# table_15_4 <- tables_2023[[16]] |>
# clean_names() |>
# glimpse()
```