---
title: "Klamath Basin Flow Gage EDA"
output: 
  html_document:
    toc_depth: 2
    theme: flatly
    code_folding: hide
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>", 
  fig.width=8, fig.height=5)
library(googleCloudStorageR)
library(tidyverse)
library(SRJPEdata)
library(lubridate)
library(sf)
library(leaflet)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", # Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87" # Isle of dogs 2 altered slightly
)
```

```{r}
sprague <- dataRetrieval::readNWISdv(11501000, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "sprague river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11501000",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11501000")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11501000")$dec_long_va # directly add lon
         )
william <- dataRetrieval::readNWISdv(11502500, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "williamson river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11502500",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11502500")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11502500")$dec_long_va # directly add lon
         )
wood <- dataRetrieval::readNWISdv(11504115, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "wood river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11504115",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11504115")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11504115")$dec_long_va # directly add lon
         )
canal <- dataRetrieval::readNWISdv(11504260, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "fourmile canal", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11504290",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11504290")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11504290")$dec_long_va # directly add lon
         )

jackson <- dataRetrieval::readNWISdv(11491470, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "jackson creek", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11491470",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11491470")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11491470")$dec_long_va # directly add lon
         ) 
irving <- dataRetrieval::readNWISdv(11491450, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "irving creek", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11491450",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11491450")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11491450")$dec_long_va # directly add lon
         ) 

random_points <- bind_rows(sprague, william, wood, canal, keno, jackson, irving)
```

```{r}
random_points_dates <- random_points |> 
  group_by(gage_number) |> 
  summarise(min_date = min(date), max_date = max(date))

unique_gage_random <- random_points |> 
  distinct(gage_number, latitude, longitude, .keep_all = TRUE) |> 
  left_join(random_points_dates, by = "gage_number")

unique_gage_random_sf <- unique_gage_random |> 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

leaflet(unique_gage_random_sf) |> 
  addTiles() |> 
  addCircleMarkers(popup = ~paste0("Stream name: ", stream, "<br>Gage Number: ", gage_number, "<br>Max Date: ", max_date, "<br>Min Date: ", min_date))
```

  - LINK RIVER AT KLAMATH FALLS, OR	11507500
  - KLAMATH RIVER BLW JOHN C.BOYLE PWRPLNT, NR KENO,OR	11510700
  - KLAMATH R NR KLAMATH CA	11530500
  - KLAMATH R A ORLEANS	11523000
  - KLAMATH RIVER AT KENO, OR	11509500
  - KLAMATH RIVER ABOVE KENO DAM, AT KENO, OR	11509370
  - KLAMATH RIVER ABV KENO DAM, AT KENO - BOTTOM	420741121554001
  - KLAMATH STRAITS DRAIN NEAR HIGHWAY 97, OR	420451121510000
  - KLAMATH STRAITS DRAIN ABOVE F-FF PUMPS, WORDEN, OR	420448121503100
  - KLAMATH RIVER AT MILLER ISLAND BOAT RAMP, OR	420853121505500
  

```{r}
klamath_fl <- dataRetrieval::readNWISdv(11507500, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11507500",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11507500")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11507500")$dec_long_va # directly add lon
         )
klamath_keno <- dataRetrieval::readNWISdv(11510700, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11510700",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11510700")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11510700")$dec_long_va # directly add lon
         )
klamath_kl <- dataRetrieval::readNWISdv(11530500, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11530500",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11530500")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11530500")$dec_long_va # directly add lon
         )
klamath_or <- dataRetrieval::readNWISdv(11523000, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11523000",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11523000")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11523000")$dec_long_va # directly add lon
         )
klamath_keno_or <- dataRetrieval::readNWISdv(11509500, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11509500",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11509500")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11509500")$dec_long_va # directly add lon
         )
klamath_ab_keno_or <- dataRetrieval::readNWISdv(11509370, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11509370",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11509370")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11509370")$dec_long_va # directly add lon
         )
klamath_ab_keno_bot <- dataRetrieval::readNWISdv(420741121554001, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", 
         gage_agency = "USGS",
         gage_number = "420741121554001",
         parameter = "temperature",
         statistic = "mean", 
         latitude = dataRetrieval::readNWISsite("420741121554001")$dec_lat_va, 
         longitude = dataRetrieval::readNWISsite("420741121554001")$dec_long_va 
         )
klamath_nr_97 <- dataRetrieval::readNWISdv(420451121510000, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", # TODO should we all it klamath straits drain?
         gage_agency = "USGS",
         gage_number = "420451121510000",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("420451121510000")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("420451121510000")$dec_long_va # directly add lon
         )
klamath_ab_pumps <- dataRetrieval::readNWISdv(420448121503100, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", # TODO should we all it klamath straits drain?
         gage_agency = "USGS",
         gage_number = "420448121503100",
         parameter = "temperature",
         statistic = "mean", 
         latitude = dataRetrieval::readNWISsite("420448121503100")$dec_lat_va, 
         longitude = dataRetrieval::readNWISsite("420448121503100")$dec_long_va 
         )
klamath_mill <- dataRetrieval::readNWISdv(420853121505500, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", 
         gage_agency = "USGS",
         gage_number = "420853121505500",
         parameter = "temperature",
         statistic = "mean", 
         latitude = dataRetrieval::readNWISsite("420853121505500")$dec_lat_va, 
         longitude = dataRetrieval::readNWISsite("420853121505500")$dec_long_va 
         )
klamath_mill_ramp <- dataRetrieval::readNWISdv(420853121505501, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "klamath river", 
         gage_agency = "USGS",
         gage_number = "420853121505501",
         parameter = "temperature",
         statistic = "mean", 
         latitude = dataRetrieval::readNWISsite("420853121505501")$dec_lat_va, 
         longitude = dataRetrieval::readNWISsite("420853121505501")$dec_long_va 
         )

mainsteam <- bind_rows(klamath_fl, klamath_keno, klamath_kl, klamath_or, klamath_keno_or, klamath_ab_keno_or, klamath_ab_keno_bot, klamath_nr_97, klamath_ab_pumps, klamath_mill, klamath_mill_ramp)

```

```{r}
# Filter to keep only one unique gage per location
all_points_dates <- mainsteam |> 
  group_by(gage_number) |> 
  summarise(min_date = min(date), max_date = max(date))

unique_gage_main <- mainsteam |> 
  distinct(gage_number, latitude, longitude, .keep_all = TRUE) |> 
  left_join(all_points_dates, by = "gage_number")

unique_gage_main_sf <- unique_gage_main |> 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

leaflet(unique_gage_main_sf) |> 
  addTiles() |> 
  addCircleMarkers(popup = ~paste0("Stream name: ", stream, "<br>Gage Number: ", gage_number, "<br>Max Date: ", max_date, "<br>Min Date: ", min_date))
```

  - TRINITY R AB NF TRINITY R NR HELENA CA	11526400
  - TRINITY R A HOOPA CA	11530000

```{r}
trinity_id <- dataRetrieval::readNWISdv(11526400, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "trinity river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11526400",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11526400")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11526400")$dec_long_va # directly add lon
         )
trinity_hoop <- dataRetrieval::readNWISdv(11530000, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "trinity river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11530000",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11530000")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11530000")$dec_long_va # directly add lon
         )

```

Other streams

  - SPRAGUE RIVER NEAR CHILOQUIN, OR	11501000
  - WILLIAMSON RIVER BLW SPRAGUE RIVER NR CHILOQUIN,OR	11502500
  - WOOD RIVER NEAR KLAMATH AGENCY, OR	11504115
  - SEVENMILE CNL AT DIKE RD BR, NR KLAMATH AGENCY, OR	11504290
  - FOURMILE CANAL NEAR KLAMATH AGENCY, OR	11504260
  - KLAMATH R AB FALL C NR COPCO CA	11511990
  - LOST RVR AT ANDERSON ROSE DVRSN DAM AT MALONE, OR	420037121334100
  - LOST RVR BLW ANDERSON ROSE DVRSN DAM AT MALONE, OR	420036121333700
  

```{r}
sprague <- dataRetrieval::readNWISdv(11501000, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "sprague river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11501000",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11501000")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11501000")$dec_long_va # directly add lon
         )

williamsom <- dataRetrieval::readNWISdv(11502500, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "williamsom river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11502500",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11502500")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11502500")$dec_long_va # directly add lon
         )

wood <- dataRetrieval::readNWISdv(11504115, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "wood river", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11504115",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11504115")$dec_lat_va, # directly add lat
         longitude = dataRetrieval::readNWISsite("11504115")$dec_long_va # directly add lon
         )

# sevenmile_cl <- dataRetrieval::readNWISdv(11504290, "00010") 
  # select(date = Date, value =  X_00010_00003) |>  # rename to value
  # as_tibble() |>
  # mutate(stream = "sevenmile canal", # add additional columns for stream, gage info, and parameter
  #        gage_agency = "USGS",
  #        gage_number = "11504290",
  #        parameter = "temperature",
  #        statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
  #        latitude = dataRetrieval::readNWISsite("11504290")$dec_lat_va, # directly add lat
  #        longitude = dataRetrieval::readNWISsite("11504290")$dec_long_va # directly add lon
  #        )

fourmile_cl <- dataRetrieval::readNWISdv(11504260, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  # rename to value
  as_tibble() |>
  mutate(stream = "", # add additional columns for stream, gage info, and parameter
         gage_agency = "USGS",
         gage_number = "11504260",
         parameter = "temperature",
         statistic = "mean", # if query returns instantaneous data then report a min, mean, and max
         latitude = dataRetrieval::readNWISsite("11504260")$dec_lat_va, 
         longitude = dataRetrieval::readNWISsite("11504260")$dec_long_va 
         )
iron_res  <- dataRetrieval::readNWISdv(11511990, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  
  as_tibble() |>
  mutate(stream = "irongate reservoir", 
         gage_agency = "USGS",
         gage_number = "11511990",
         parameter = "temperature",
         statistic = "mean", 
         latitude = dataRetrieval::readNWISsite("11511990")$dec_lat_va, 
         longitude = dataRetrieval::readNWISsite("11511990")$dec_long_va 
         )
lost_rvr  <- dataRetrieval::readNWISdv(420037121334100, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  
  as_tibble() |>
  mutate(stream = "lost river", 
         gage_agency = "USGS",
         gage_number = "420037121334100",
         parameter = "temperature",
         statistic = "mean", 
         latitude = dataRetrieval::readNWISsite("420037121334100")$dec_lat_va, 
         longitude = dataRetrieval::readNWISsite("420037121334100")$dec_long_va 
         )
lost_rvr_and  <- dataRetrieval::readNWISdv(420036121333700, "00010") |> 
  select(date = Date, value =  X_00010_00003) |>  
  as_tibble() |>
  mutate(stream = "lost river", 
         gage_agency = "USGS",
         gage_number = "420036121333700",
         parameter = "temperature",
         statistic = "mean", 
         latitude = dataRetrieval::readNWISsite("420036121333700")$dec_lat_va, 
         longitude = dataRetrieval::readNWISsite("420036121333700")$dec_long_va 
         )

other_temp_points <- bind_rows(sprague, williamsom, wood, fourmile_cl, lost_rvr, lost_rvr_and)
```

```{r}
# Filter to keep only one unique gage per location
all_points_dates <- other_temp_points |> 
  group_by(gage_number) |> 
  summarise(min_date = min(date), max_date = max(date))

unique_gage_data <- other_temp_points |> 
  distinct(gage_number, latitude, longitude, .keep_all = TRUE) |> 
  left_join(all_points_dates, by = "gage_number")

unique_gage_sf <- unique_gage_data |> 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

leaflet(unique_gage_sf) |> 
  addTiles() |> 
  addCircleMarkers(popup = ~paste0("Stream name: ", stream, "<br>Gage Number: ", gage_number, "<br>Max Date: ", max_date, "<br>Min Date: ", min_date))
```


### Map Plot Klamath USGS Gages

```{r}
# Filter to keep only one unique gage per location
all_points_dates <- all_points |> 
  group_by(gage_number) |> 
  summarise(min_date = min(date), max_date = max(date))

unique_gage_data <- all_points |> 
  distinct(gage_number, latitude, longitude, .keep_all = TRUE) |> 
  left_join(all_points_dates, by = "gage_number")

unique_gage_sf <- unique_gage_data |> 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

leaflet(unique_gage_sf) |> 
  addTiles() |> 
  addCircleMarkers(popup = ~paste0("Stream name: ", stream, "<br>Gage Number: ", gage_number, "<br>Max Date: ", max_date, "<br>Min Date: ", min_date))
```
