---
title: "Modeled Fall Chinook Population Data"
author: "Ashley Vizek"
date: "March 2025"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(scales)
library(readxl)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", "#D67236",# Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87", # Isle of dogs 2 altered slightly
                  "#F3DF6C", "#CEAB07", "#D5D5D3", "#24281A", # Moonriese 1, 
                  "#798E87", "#C27D38", "#CCC591", "#29211F", # moonrise 2
                  "#85D4E3", "#F4B5BD", "#9C964A", "#CDC08C", "#FAD77B" ) # moonrise 3 

colors_two <- c("#D8A499", "#74A089")

# define AWS data bucket
# note that you need to set up access keys in R environ
klamath_project_board <- pins::board_s3(
  bucket="klamath-sdm",
  access_key=Sys.getenv("aws_access_key_id"),
  secret_access_key=Sys.getenv("secret_access_key_id"),
  session_token = Sys.getenv("session_token_id"),
  region = "us-east-1"
)
```

The purpose of this markdown is to give an overview of the dataset including specific data types, temporal coverage, and geographic coverage. This document is helpful in developing schemas to bring together disparate datasets and exploring data limitations.

# Overview

**Data type:** Fall Chinook escapement on the mainstem Klamath River

These data were scraped from a US Fish and Wildlife Service Technical Report using the Tabula software where you import a PDF and select a table to export: 

[Gough, S. A., C. Z. Romberger, and N. A. Som. 2018. Fall Chinook Salmon Run Characteristics and Escapement in the Mainstem Klamath River below Iron Gate Dam, 2017. U.S. Fish and Wildlife Service. Arcata Fish and Wildlife Office, Arcata Fisheries Data Series Report Number DS 2018–58, Arcata, California](https://www.fws.gov/sites/default/files/documents/2017%20klamath%20spawn%20survey%20report%202017%20FINAL1.pdf)

## Data Available {.tabset} 

**Annual adult abundance**

- Higher adult abundance in Scott 
- Larger variation in abundance across years in Scott
- Longer time series in Shasta

### Klamath Fall Chinook escapement

```{r, include=FALSE}
escapement_raw <- read_csv(here::here("data-raw", "fisheries", "modeled", "fall_chinook_escapement.csv"))

escapement_processed <- escapement_raw |> 
  rename(julian_year = Year,
         estimation_method = Estimator) |> 
  separate(`Lower Upper`, into = c("lower_bounds_estimate", "upper_bounds_estimate"), sep = " ") |> 
  mutate(estimate = gsub(",","",estimate),
         estimate = gsub(" ", "",estimate),
         estimate = as.numeric(estimate),
         lower_bounds_estimate = as.numeric(gsub(",","",lower_bounds_estimate)),
         upper_bounds_estimate = as.numeric(gsub(",","",upper_bounds_estimate)),
         estimation_method = tolower(estimation_method),
         species = "fall chinook",
         stream = "klamath")
```

 - Title: Annual escapement estimates for Klamath River fall Chinook
 - Spatial extent: mainstem Klamath River
 - Description: Adult fall Chinook Salmon carcasses and redds were surveyed on the mainstem Klamath River, from Iron Gate Dam to Wingate Bar during the 2017 spawning season to estimate annual escapement and characterize the age and sex composition and spawning success of the run. Surveys were conducted over 9 weeks, from October 11 to December 6. Using postmortem mark–recapture methods and a hierarchical latent variables model between Iron Gate Dam and the confluence with the Shasta River, the estimated spawning escapement for this 21.6-km section of the mainstem Klamath River was 4,740 fish.
 - Temporal extent and resolution: 2001-2017, annual
 - Fields included: `r colnames(escapement_raw)`
 - Name of data file: fall_chinook_escapement.csv
 
**Data processing implications:**

- Need to split the lower and upper bounds
- Tranform character to numeric
- Clean up the column names

**Questions:**
- What is going on in the years with high uncertainty


**Action items:**
- Try to find a more updated table with data through present
- Try to find if these data are available in other locations

#### Data summary

This section may (or may not) explore every variable in the dataset. It will include high level summaries of the data to communicate the primary story of the data.

```{r, echo = FALSE}
escapement_processed |> 
  ggplot(aes(x = julian_year, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_bounds_estimate, ymax = upper_bounds_estimate), width = 0.2) + 
  theme_bw() +
  labs(x = "",
       y = "klamath river fall chinook escapement estimate")

```


## Save data

Data files are stored on AWS

```{r, include = F}
klamath_project_board |> pins::pin_write(escapement_processed,
                               type = "csv",
                               title = "fall_chinook_escapement")
```