---
title: "Modeled Upper Klamath Lake Sucker Survival Data"
author: "Ashley Vizek"
date: "March 2025"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(scales)
library(readxl)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", "#D67236",# Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87", # Isle of dogs 2 altered slightly
                  "#F3DF6C", "#CEAB07", "#D5D5D3", "#24281A", # Moonriese 1, 
                  "#798E87", "#C27D38", "#CCC591", "#29211F", # moonrise 2
                  "#85D4E3", "#F4B5BD", "#9C964A", "#CDC08C", "#FAD77B" ) # moonrise 3 

colors_3 <- c("#D8A499", "#74A089", "#9A8822")

# define AWS data bucket
# note that you need to set up access keys in R environ
klamath_project_board <- pins::board_s3(
  bucket="klamath-sdm",
  access_key=Sys.getenv("aws_access_key_id"),
  secret_access_key=Sys.getenv("secret_access_key_id"),
  session_token = Sys.getenv("session_token_id"),
  region = "us-east-1"
)
```

The purpose of this markdown is to give an overview of the dataset including specific data types, temporal coverage, and geographic coverage. This document is helpful in developing schemas to bring together disparate datasets and exploring data limitations.

# Overview

**Data type:** Adult survival estimates of Lost River and Shortnose suckers in Upper Klamath Lake

These data were scraped from a USGS Technical Report using the Tabula software where you import a PDF and select a table to export: 

[Hewitt, D.A., Janney, E.C., Hayes, B.S., and Harris, A.C., 2018, Status and trends of adult Lost River (Deltistes luxatus) and shortnose (Chasmistes brevirostris) sucker populations in Upper Klamath Lake, Oregon, 2017: U.S. Geological Survey Open-File Report 2018-1064, 31 p., https://doi.org/10.3133/ofr20181064.](https://pubs.usgs.gov/of/2018/1064/ofr20181064.pdf)

The raw data associated with this report are available; however, we decided to focus on the modeled output first. Ultimately we envision pulling in raw survival data from the KBFC PIT tag database.

## Data Available {.tabset} 

- Species: Lost River (lake and river spawning) and Shortnose
- Lifestage: Adult
- Parameters: apparent survival, seniority probability, annual population rate of change

### Adult sucker survival

```{r, include=FALSE}
survival_raw <- read_csv(here::here("data-raw", "fisheries", "modeled", "sucker_survival.csv"))
```

 - Title: Annual adult survival estimates for Lost River and Shortnose suckers in Upper Klamath Lake
 - Spatial extent: Upper Klamath Lake
 - Description: 
 - Temporal extent and resolution: 1999-2016, annual
 - Fields included: `r colnames(survival_raw)`
 - Name of data file: sucker_survival.csv
 
**Data processing implications:**

- Remove estimates labelled "B" or "C"
- Split the CI into the low/high bounds
- We could add this to other modelled data

**Questions:**

- I thought the population of declining but population change shows small increase


**Action items:**


#### Data summary

This section may (or may not) explore every variable in the dataset. It will include high level summaries of the data to communicate the primary story of the data.

```{r, echo = FALSE}
survival_raw |> 
  filter(!apparent_survival_estimate %in% c("B","C")) |> 
  separate(apparent_survival_CI, into = c("low_ci", "high_ci"), sep = "-") |> 
  mutate(apparent_survival_estimate = as.numeric(apparent_survival_estimate),
         low_ci = as.numeric(low_ci),
         high_ci = as.numeric(high_ci)) |> 
  ggplot(aes(x = year, y = apparent_survival_estimate, colour = population)) +
  geom_point() +
  geom_errorbar(aes(ymin = low_ci, ymax = high_ci), width = 0.2) + 
  scale_color_manual(values = colors_3) +
  theme_bw() +
  facet_wrap(~sex) +
  labs(x = "",
       y = "apparent survival",
       color = "")

survival_raw |> 
  filter(!seniority_probability_estimate %in% c("B","C")) |> 
  separate(seniority_probability_CI, into = c("low_ci", "high_ci"), sep = "-") |> 
  mutate(seniority_probability_estimate = as.numeric(seniority_probability_estimate),
         low_ci = as.numeric(low_ci),
         high_ci = as.numeric(high_ci)) |> 
  ggplot(aes(x = year, y = seniority_probability_estimate, colour = population)) +
  geom_point() +
  geom_errorbar(aes(ymin = low_ci, ymax = high_ci), width = 0.2) + 
  scale_color_manual(values = colors_3) +
  theme_bw() +
  facet_wrap(~sex) +
  labs(x = "",
       y = "seniority probability",
       color = "")

survival_raw |> 
  filter(!annual_population_rate_of_change_estimate %in% c("B","C")) |> 
  separate(annual_population_rate_of_change_CI, into = c("low_ci", "high_ci"), sep = "-") |> 
  mutate(annual_population_rate_of_change_estimate = as.numeric(annual_population_rate_of_change_estimate),
         low_ci = as.numeric(low_ci),
         high_ci = as.numeric(high_ci)) |> 
  ggplot(aes(x = year, y = annual_population_rate_of_change_estimate, colour = population)) +
  geom_point() +
  geom_errorbar(aes(ymin = low_ci, ymax = high_ci), width = 0.2) + 
  scale_color_manual(values = colors_3) +
  theme_bw() +
  facet_wrap(~sex) +
  labs(x = "",
       y = "population rate of change",
       color = "")

```

## Save data

Data files are stored on AWS

```{r, include = F}
klamath_project_board |> pins::pin_write(survival_raw,
                               type = "csv",
                               title = "sucker_adult_survival_model")
```