---
title: "Modeled Population Data from CDFW"
author: "Ashley Vizek"
date: "March 2025"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(scales)
library(readxl)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", "#D67236",# Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87", # Isle of dogs 2 altered slightly
                  "#F3DF6C", "#CEAB07", "#D5D5D3", "#24281A", # Moonriese 1, 
                  "#798E87", "#C27D38", "#CCC591", "#29211F", # moonrise 2
                  "#85D4E3", "#F4B5BD", "#9C964A", "#CDC08C", "#FAD77B" ) # moonrise 3 

colors_two <- c("#D8A499", "#74A089")
colors_five <- c("#9A8822","#5B1A18","#74A089","#85D4E3","#D8A499")

# define AWS data bucket
# note that you need to set up access keys in R environ
klamath_project_board <- pins::board_s3(
  bucket="klamath-sdm",
  access_key=Sys.getenv("aws_access_key_id"),
  secret_access_key=Sys.getenv("secret_access_key_id"),
  session_token = Sys.getenv("session_token_id"),
  region = "us-east-1"
)
```

The purpose of this markdown is to give an overview of the dataset including specific data types, temporal coverage, and geographic coverage. This document is helpful in developing schemas to bring together disparate datasets and exploring data limitations.

# Overview

**Data type:** Salmonid Population Data

These data were downloaded from California Department of Fish and Wildlife (CDFW)
[document library](https://www.nrm.dfg.ca.gov/documents/ContextDocs.aspx?cat=Fisheries--AnadromousSalmonidPopulationMonitoring)

## Data Available {.tabset} 

This dataset spans watersheds throughout California. We filtered to Trinity River,
Scott River, Shasta River, Lower Klamath, and Klamath River.

Populations included: Bogus Creek, Lower Klamath, Scott River, Shasta River, Trinity River

Species included: coho, steelhead, chinook

Runs included: winter, fall, spring

Lifestage included: adult, YOY, age 1+, age 2+, adult and subadult, smolt

### Salmonid Population Data from CDFW

```{r, include=FALSE}
cdfw_population_raw <- read_xlsx(here::here("data-raw", "fisheries", "modeled", "Salmonid_Population_Monitoring_Data_CMPv2023.xlsx"), sheet = "Population Data")

klamath_cdfw_population_raw <- cdfw_population_raw |> 
  filter(Watershed %in% c("Trinity River", "Scott River", "Shasta River", "Lower Klamath","Klamath River"))

# escapement_processed <- escapement_raw |> 
#   rename(julian_year = Year,
#          estimation_method = Estimator) |> 
#   separate(`Lower Upper`, into = c("lower_bounds_estimate", "upper_bounds_estimate"), sep = " ") |> 
#   mutate(estimate = gsub(",","",estimate),
#          estimate = gsub(" ", "",estimate),
#          estimate = as.numeric(estimate),
#          lower_bounds_estimate = as.numeric(gsub(",","",lower_bounds_estimate)),
#          upper_bounds_estimate = as.numeric(gsub(",","",upper_bounds_estimate)),
#          estimation_method = tolower(estimation_method),
#          species = "fall chinook",
#          stream = "klamath")

klamath_cdfw_population_processed <- klamath_cdfw_population_raw |> 
  janitor::clean_names() |> 
  rename(stream = population,
         lifestage = life_stage) |> 
    mutate(species = ifelse(!is.na(run_designation), tolower(paste0(run_designation, " ", species)), tolower(species)),
           julian_year = as.numeric(ifelse(!is.na(brood_year), brood_year, survey_season)),
           stream = tolower(stream),
           lifestage = tolower(lifestage),
           estimate_type = tolower(metric),
           estimation_method = tolower(estimation_method),
           sex = NA,
           estimate = value,
           confidence_interval = 95,
           lower_bounds_estimate = x95_lower_ci,
           upper_bounds_estimate = x95_upper_ci,
           is_complete_estimate = ifelse(full_population_estimate == "N", F, T)) |> 
  select(julian_year, stream, species, lifestage, sex, estimate_type, estimate, confidence_interval,
         lower_bounds_estimate, upper_bounds_estimate, estimation_method, is_complete_estimate)
```

 - Title: Salmonid population data from CDFW
 - Spatial extent: Bogus Creek, Lower Klamath, Scott River, Shasta River, Trinity River
 - Description: The California Monitoring Plan (CMP) salmonid monitoring data provides a summary of salmonid population metrics and corresponding authors or sources. This information was updated during the fall and winter of 2023 and includes select data from 1978 through 2023 but is not considered comprehensive. The population metrics summarized are focused on the viable salmonid population (VSP) key characteristics: abundance, productivity, and spatial structure. It is not recommended for users to extract data and make inferences without appropriately reviewing the associated source document(s). This dataset does not include data for all existing or past monitoring programs. The term “steelhead” is used in the Species field but does not confirm an anadromous life history. “Steelhead” values and estimates should be conservatively interpreted as Oncorhynchus mykiss (Steelhead or rainbow trout) and further investigated using the Source field in each record. Please read Salmonid Population Monitoring Data Description_CMP2021 for more information.
 - Temporal extent and resolution: 1978-2023, annual (varies by location and species)
 - Fields included: `r colnames(klamath_cdfw_population_raw)`
 - Name of data file: Salmonid_Population_Monitoring_Data_CMPv2023.xlsx
 
**Data processing implications:**

- column names to snake case
- transform to format outlined in schema

**Questions:**
- how are these data different than estimates gathered previously (e.g. scott and shasta coho): they are the same but these data are more up to date so use these
- where do these data come from? where are the methods?

**Action items:**


#### Data summary

This section may (or may not) explore every variable in the dataset. It will include high level summaries of the data to communicate the primary story of the data.

```{r, echo = FALSE}

# coho
# spring chinook
# fall chinook
# winter steelhead
# steelhead

klamath_cdfw_population_processed |> 
  filter(species == "coho salmon") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "coho",
       color = "")

klamath_cdfw_population_processed |> 
  filter(species == "spring chinook salmon") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "spring chinook",
       color = "")

klamath_cdfw_population_processed |> 
  filter(species == "fall chinook salmon") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "fall chinook",
       color = "")

klamath_cdfw_population_processed |> 
  filter(species == "winter steelhead") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "winter steelhead",
       color = "")

klamath_cdfw_population_processed |> 
  filter(species == "steelhead") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "steelhead",
       color = "")

```


## Save data

Data files are stored on AWS

```{r, include = F}
klamath_project_board |> pins::pin_write(klamath_cdfw_population_processed,
                               type = "csv",
                               title = "klamath_cdfw_population")
```