---
title: "Modeled Fisheries Data"
author: "Ashley Vizek"
date: "March 2025"
output: rmarkdown::github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(scales)
library(readxl)

colors_full <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                  "#899DA4", "#C93312", "#DC863B", # royal 1 (- 3)
                  "#F1BB7B", "#FD6467", "#5B1A18", "#D67236",# Grand Budapest 1 (-4)
                  "#D8B70A", "#02401B", "#A2A475", # Cavalcanti 1
                  "#E6A0C4", "#C6CDF7", "#D8A499", "#7294D4", #Grand Budapest 2
                  "#9986A5", "#EAD3BF", "#AA9486", "#B6854D", "#798E87", # Isle of dogs 2 altered slightly
                  "#F3DF6C", "#CEAB07", "#D5D5D3", "#24281A", # Moonriese 1, 
                  "#798E87", "#C27D38", "#CCC591", "#29211F", # moonrise 2
                  "#85D4E3", "#F4B5BD", "#9C964A", "#CDC08C", "#FAD77B" ) # moonrise 3 

colors_two <- c("#D8A499", "#74A089")
colors_3 <- c("#D8A499", "#74A089", "#9A8822")
colors_five <- c("#9A8822","#5B1A18","#74A089","#85D4E3","#D8A499")

# define AWS data bucket
# note that you need to set up access keys in R environ
klamath_project_board <- pins::board_s3(
  bucket="klamath-sdm",
  access_key=Sys.getenv("aws_access_key_id"),
  secret_access_key=Sys.getenv("secret_access_key_id"),
  session_token = Sys.getenv("session_token_id"),
  region = "us-east-1"
)
```

The purpose of this markdown is to give an overview of the dataset including specific data types, temporal coverage, and geographic coverage. This document is helpful in developing schemas to bring together disparate datasets and exploring data limitations.

# Overview {.tabset}

**Data type:** modeled fisheries data

These data will ultimately be part of the [klamathFishData](https://github.com/Klamath-SDM/klamathFishData) R package. This document summarizes data from a number of sources.

## Salmonid Population Data from CDFW

This dataset spans watersheds throughout California. We filtered to Trinity River,
Scott River, Shasta River, Lower Klamath, and Klamath River.

- Populations included: Bogus Creek, Lower Klamath, Scott River, Shasta River, Trinity River
- Species included: coho, steelhead, chinook
- Runs included: winter, fall, spring
- Lifestage included: adult, YOY, age 1+, age 2+, adult and subadult, smolt

Note that this is a synthesis spreadsheet and not the original source of data. 
- Coho data for the Scott and Shasta can also be found:
[Manhard, C. V., N. A. Som, R. W. Perry, J. R. Faukner, and T. Soto. 2018.  Estimating freshwater productivity, overwinter survival, and migration patterns of Klamath River Coho Salmon.  U.S. Fish and Wildlife Service.  Arcata Fish and Wildlife Office, Arcata Fisheries Technical Report Number TR 2018-33, Arcata, California.](https://www.fws.gov/sites/default/files/documents/EstimatingFreshwaterProductivityOverwinterSurvivalandMigrationPatternsofKlamathRiverCohoSalmon.pdf)


```{r, include=FALSE}
cdfw_population_raw <- read_xlsx(here::here("data-raw", "fisheries", "modeled", "Salmonid_Population_Monitoring_Data_CMPv2023.xlsx"), sheet = "Population Data")

klamath_cdfw_population_raw <- cdfw_population_raw |> 
  filter(Watershed %in% c("Trinity River", "Scott River", "Shasta River", "Lower Klamath","Klamath River"))

klamath_cdfw_population_processed <- klamath_cdfw_population_raw |> 
  janitor::clean_names() |> 
  rename(stream = population,
         lifestage = life_stage) |> 
    mutate(species = ifelse(!is.na(run_designation), tolower(paste0(run_designation, " ", species)), tolower(species)),
           julian_year = as.numeric(ifelse(!is.na(brood_year), brood_year, survey_season)),
           stream = tolower(stream),
           lifestage = tolower(lifestage),
           estimate_type = tolower(metric),
           estimation_method = tolower(estimation_method),
           sex = NA,
           estimate = value,
           confidence_interval = 95,
           lower_bounds_estimate = x95_lower_ci,
           upper_bounds_estimate = x95_upper_ci,
           is_complete_estimate = ifelse(full_population_estimate == "N", F, T)) |> 
  select(julian_year, stream, species, lifestage, sex, estimate_type, estimate, confidence_interval,
         lower_bounds_estimate, upper_bounds_estimate, estimation_method, is_complete_estimate)
```

 - Title: Salmonid population data from CDFW
 - Spatial extent: Bogus Creek, Lower Klamath, Scott River, Shasta River, Trinity River
 - Description: The California Monitoring Plan (CMP) salmonid monitoring data provides a summary of salmonid population metrics and corresponding authors or sources. This information was updated during the fall and winter of 2023 and includes select data from 1978 through 2023 but is not considered comprehensive. The population metrics summarized are focused on the viable salmonid population (VSP) key characteristics: abundance, productivity, and spatial structure. It is not recommended for users to extract data and make inferences without appropriately reviewing the associated source document(s). This dataset does not include data for all existing or past monitoring programs. The term “steelhead” is used in the Species field but does not confirm an anadromous life history. “Steelhead” values and estimates should be conservatively interpreted as Oncorhynchus mykiss (Steelhead or rainbow trout) and further investigated using the Source field in each record. Please read Salmonid Population Monitoring Data Description_CMP2021 for more information.
 - Temporal extent and resolution: 1978-2023, annual (varies by location and species)
 - Fields included: `r colnames(klamath_cdfw_population_raw)`
 - Name of data file: Salmonid_Population_Monitoring_Data_CMPv2023.xlsx
 - Source: These data were downloaded from California Department of Fish and Wildlife (CDFW)
[document library](https://www.nrm.dfg.ca.gov/documents/ContextDocs.aspx?cat=Fisheries--AnadromousSalmonidPopulationMonitoring)
 
**Data processing implications:**

- column names to snake case
- transform to format outlined in schema

**Questions:**

- how are these data different than estimates gathered previously (e.g. scott and shasta coho): they are the same but these data are more up to date so use these
- where do these data come from? where are the methods?

**Action items:**


### Data summary

This section may (or may not) explore every variable in the dataset. It will include high level summaries of the data to communicate the primary story of the data.

```{r, echo = FALSE}

# coho
# spring chinook
# fall chinook
# winter steelhead
# steelhead

klamath_cdfw_population_processed |> 
  filter(species == "coho salmon") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "coho",
       color = "")

klamath_cdfw_population_processed |> 
  filter(species == "spring chinook salmon") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "spring chinook",
       color = "")

klamath_cdfw_population_processed |> 
  filter(species == "fall chinook salmon") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "fall chinook",
       color = "")

klamath_cdfw_population_processed |> 
  filter(species == "winter steelhead") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "winter steelhead",
       color = "")

klamath_cdfw_population_processed |> 
  filter(species == "steelhead") |> 
  ggplot(aes(x = julian_year, y = estimate, color = stream)) +
  geom_line() +
  scale_color_manual(values = colors_five) +
  facet_wrap(~lifestage, scales = "free_y") +
  theme_bw() +
  labs(x = "",
       y = "Abundance estimate",
       title = "steelhead",
       color = "")

```

## Klamath Fall Chinook escapement

```{r, include=FALSE}
escapement_raw <- read_csv(here::here("data-raw", "fisheries", "modeled", "fall_chinook_escapement.csv"))

klamath_mainstem_fall_chinook_escapement <- escapement_raw |> 
  rename(julian_year = Year,
         estimation_method = Estimator) |> 
  separate(`Lower Upper`, into = c("lower_bounds_estimate", "upper_bounds_estimate"), sep = " ") |> 
  mutate(estimate = gsub(",","",estimate),
         estimate = gsub(" ", "",estimate),
         estimate = as.numeric(estimate),
         lower_bounds_estimate = as.numeric(gsub(",","",lower_bounds_estimate)),
         upper_bounds_estimate = as.numeric(gsub(",","",upper_bounds_estimate)),
         estimation_method = tolower(estimation_method),
         species = "fall chinook",
         stream = "klamath")
```

 - Title: Annual escapement estimates for Klamath River fall Chinook
 - Spatial extent: mainstem Klamath River
 - Description: Adult fall Chinook Salmon carcasses and redds were surveyed on the mainstem Klamath River, from Iron Gate Dam to Wingate Bar during the 2017 spawning season to estimate annual escapement and characterize the age and sex composition and spawning success of the run. Surveys were conducted over 9 weeks, from October 11 to December 6. Using postmortem mark–recapture methods and a hierarchical latent variables model between Iron Gate Dam and the confluence with the Shasta River, the estimated spawning escapement for this 21.6-km section of the mainstem Klamath River was 4,740 fish.
 - Temporal extent and resolution: 2001-2017, annual
 - Fields included: `r colnames(escapement_raw)`
 - Name of data file: fall_chinook_escapement.csv
 - Source: These data were scraped from a US Fish and Wildlife Service Technical Report using the Tabula software where you import a PDF and select a table to export: 

[Gough, S. A., C. Z. Romberger, and N. A. Som. 2018. Fall Chinook Salmon Run Characteristics and Escapement in the Mainstem Klamath River below Iron Gate Dam, 2017. U.S. Fish and Wildlife Service. Arcata Fish and Wildlife Office, Arcata Fisheries Data Series Report Number DS 2018–58, Arcata, California](https://www.fws.gov/sites/default/files/documents/2017%20klamath%20spawn%20survey%20report%202017%20FINAL1.pdf)
 
**Data processing implications:**

- Need to split the lower and upper bounds
- Tranform character to numeric
- Clean up the column names

**Questions:**
- What is going on in the years with high uncertainty


**Action items:**
- Try to find a more updated table with data through present
- Try to find if these data are available in other locations

### Data summary

This section may (or may not) explore every variable in the dataset. It will include high level summaries of the data to communicate the primary story of the data.

```{r, echo = FALSE}
klamath_mainstem_fall_chinook_escapement |> 
  ggplot(aes(x = julian_year, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_bounds_estimate, ymax = upper_bounds_estimate), width = 0.2) + 
  theme_bw() +
  labs(x = "",
       y = "klamath river fall chinook escapement estimate")

```

## Adult sucker survival

```{r, include=FALSE}
sucker_survival_model <- read_csv(here::here("data-raw", "fisheries", "modeled", "sucker_survival.csv"))
```

- Title: Annual adult survival estimates for Lost River and Shortnose suckers in Upper Klamath Lake
- Spatial extent: Upper Klamath Lake
- Description: 
- Temporal extent and resolution: 1999-2016, annual
- Fields included: `r colnames(sucker_survival_model)`
- Name of data file: sucker_survival.csv
- Source: These data were scraped from a USGS Technical Report using the Tabula software where you import a PDF and select a table to export: 

[Hewitt, D.A., Janney, E.C., Hayes, B.S., and Harris, A.C., 2018, Status and trends of adult Lost River (Deltistes luxatus) and shortnose (Chasmistes brevirostris) sucker populations in Upper Klamath Lake, Oregon, 2017: U.S. Geological Survey Open-File Report 2018-1064, 31 p., https://doi.org/10.3133/ofr20181064.](https://pubs.usgs.gov/of/2018/1064/ofr20181064.pdf)

The raw data associated with this report are available; however, we decided to focus on the modeled output first. Ultimately we envision pulling in raw survival data from the KBFC PIT tag database.
 
**Data processing implications:**

- Remove estimates labelled "B" or "C"
- Split the CI into the low/high bounds
- We could add this to other modelled data

**Questions:**

- I thought the population of declining but population change shows small increase

**Action items:**


### Data summary

This section may (or may not) explore every variable in the dataset. It will include high level summaries of the data to communicate the primary story of the data.

```{r, echo = FALSE}
sucker_survival_model |> 
  filter(!apparent_survival_estimate %in% c("B","C")) |> 
  separate(apparent_survival_CI, into = c("low_ci", "high_ci"), sep = "-") |> 
  mutate(apparent_survival_estimate = as.numeric(apparent_survival_estimate),
         low_ci = as.numeric(low_ci),
         high_ci = as.numeric(high_ci)) |> 
  ggplot(aes(x = year, y = apparent_survival_estimate, colour = population)) +
  geom_point() +
  geom_errorbar(aes(ymin = low_ci, ymax = high_ci), width = 0.2) + 
  scale_color_manual(values = colors_3) +
  theme_bw() +
  facet_wrap(~sex) +
  labs(x = "",
       y = "apparent survival",
       color = "")

sucker_survival_model |> 
  filter(!seniority_probability_estimate %in% c("B","C")) |> 
  separate(seniority_probability_CI, into = c("low_ci", "high_ci"), sep = "-") |> 
  mutate(seniority_probability_estimate = as.numeric(seniority_probability_estimate),
         low_ci = as.numeric(low_ci),
         high_ci = as.numeric(high_ci)) |> 
  ggplot(aes(x = year, y = seniority_probability_estimate, colour = population)) +
  geom_point() +
  geom_errorbar(aes(ymin = low_ci, ymax = high_ci), width = 0.2) + 
  scale_color_manual(values = colors_3) +
  theme_bw() +
  facet_wrap(~sex) +
  labs(x = "",
       y = "seniority probability",
       color = "")

sucker_survival_model |> 
  filter(!annual_population_rate_of_change_estimate %in% c("B","C")) |> 
  separate(annual_population_rate_of_change_CI, into = c("low_ci", "high_ci"), sep = "-") |> 
  mutate(annual_population_rate_of_change_estimate = as.numeric(annual_population_rate_of_change_estimate),
         low_ci = as.numeric(low_ci),
         high_ci = as.numeric(high_ci)) |> 
  ggplot(aes(x = year, y = annual_population_rate_of_change_estimate, colour = population)) +
  geom_point() +
  geom_errorbar(aes(ymin = low_ci, ymax = high_ci), width = 0.2) + 
  scale_color_manual(values = colors_3) +
  theme_bw() +
  facet_wrap(~sex) +
  labs(x = "",
       y = "population rate of change",
       color = "")

```

## Save data

Data files are stored on AWS

```{r, include = F}
klamath_project_board |> pins::pin_write(klamath_cdfw_population_processed,
                               type = "csv",
                               title = "klamath_cdfw_population")
klamath_project_board |> pins::pin_write(klamath_mainstem_fall_chinook_escapement,
                               type = "csv",
                               title = "klamath_mainstem_fall_chinook_escapement")

klamath_project_board |> pins::pin_write(sucker_survival_model,
                               type = "csv",
                               title = "sucker_adult_survival_model")
```